\chapter{Free Love, Sexism and All That}
\chapterauthor{Robert Anton Wilson}

\blockquote{I have said what I have said; I have not said what I have not said.
\par\begin{flushright} --- Count Alfred Korzybski \end{flushright}
}
\blockquote{Lie down on the floor and keep calm.
\par\begin{flushright} --- John Herbert Dillinger \end{flushright}
}

Anarchism derives from the Greek \emph{an-archos}, without rule, and thereby implies individual freedom. To a naive observer, this would seem to necessarily include sexual freedom; and, indeed, the majority of anarchists, until about 1968 e.v., believed that an anarchist society would be a sexually free society. Such pioneer libertarians as Josiah Warren, Ezra Heywood, Stephen Pearl Andrews, Emma Goldman, etc. vigorously joined the socialist-feminist Victoria Woodhull in defending the ideology called ``free love," i.e. sexual self-determination. The thought ``individual freedom includes sexual freedom" was so ubiquitous at the beginnings of modern radicalism, in fact, that it actually was espoused by Frederick Engels (although not, of course, by Karl Marx) and spread far beyond anarchism into the mainstream of revolutionary ideology.\\
In the past few years we have learned that this seemingly self-evident inclusion of sexual freedom withing personal freedom is not self-evident at all. Many believe that an anarchist society would not, should not and could not allow free love. Sexual self-determination, we are soberly assured, is a form of sexism, and sexism is tyranny; ergo, anarchist society must set limits on freedom just like any other society.\\
I propose to examine this argument and demonstrate its total falsity. The classic declaration of free love --- Victoria Woodhull's famous outburst, ``I have an inalienable, constitutional, and natural right to love whom I may, to love as long or as short a period as I can, to change that love every day if I please! And with that right neither you nor any law you can frame have any right to interfere" --- is, I will show, the only sexual philosophy consistent with anarchist theory.\\
There are nine forms of anti-free-love ideology currently espoused in anarchist publications; I will discuss them in ascending order, from the most amoeboid tropism-level up to the mental or philosophical level.\\
First there is the anti-free-love \emph{reflex}. This occurs in persons who have never learned to use the higher neurological circuits of the cortex; it exemplifies the classic reflex arc beloved in behaviorist theory. The phototropic moth flies straight into the torch, and dies; the conditioned dog in Dr. Pavlov's chamber of horrors drools when the dinner-bell rings; many contemporary radicals just as automatically froth at the mouth or burst into angry speech (glandular spasms defending ideological ``territories" based on old mammal reflexes) when ``free love" is mentioned. Often, there is no thought involved at all; such persons merely know that it is fashionable to have this reflex at this time.\\
The pernicious prevalence of this type of robotism is largely die to the influence of Franz Fanon, whose \emph{apologia} for revolutionary rage has been widely used and abused as a rationale for retaining into adult life the trigger-reflexes of human infants or even of pekinese dogs. This ``existentialist" politics ---
\begin{addmargin}{2cm}
When in danger or in doubt\\
Run in circles, scream and shout
\end{addmargin}
--- wishes to reduce human neurology to that of, say, the rabbit's fear-immobility mechanism. It is worth noting that this freeze-reflex works well enough when the rabbit is in tall grass or shrubs, where statue-like stillness may render him invisible to predators. When the rabbit blindly follows this reflex on the highway, he usually gets run over. Evolution has formed the higher nervous circuits as monitors or feedbacks, correcting this ``blind' robotism in the lower circuits.\\
True freedom, then, begins with neurological freedom --- detachment from the mechanistic emotional-glandular reactions. In this ``cortical delay" (as Korzybski called it) so-called ``free will" begins --- not the gift of the gods, as theologians imagine, but the result of self-work and self-reprogramming. Lacking such integral meta-programming we continue to repeat the old imprints of territorial rage-fear going back to the Carboniferous amphibians, all still coded in the midbrain and ready for operation whenever the higher circuits are deliberately bypassed.\\
The anti-free-love reflex, then, is not merely anti-logical but anti-neurological. It uses the oldest, most primitive part of the brain and turns off the more recent circuits of the past hundred million years, especially the cortical development of the hominid stock, since 100,000 B.C.E.\\
The second variety of anti-free-love ideology is based on \emph{argumentum ad hominem} and/or psychological ``analysis." Thus, when the free love issue was debated in the \emph{SRAF Bulletin}, the opponents of sexual self-determination seldom argued the theory itself but merely reiterated endlessly that exponents of free love are (a) sexists (b) male chauvinists (c) compulsive masturbators or (d) otherwise contemptible, i.e. pissy, shitty, nasty, etc. (This is not a caricature of their writings but an accurate report.)\\
This is not a substantive argument. It might be proved conceivably, that all SRAF writers who support free love are somehow vicious characters, but this would not itself demonstrate the falsity of their position. Similarly, if all Darwinians were shown to be addicted to bestiality, this would \emph{not} mean that evolution is false and we must accept \emph{Genesis}; if all modern geographers are boot-fetishists we need not then accept the Flat Earth theory; etc.\\
The third anti-free-love argument is the claim that discussion of the subject is itself invasive or coercive; that is, to debate the issue is to cause acute psychological damage to certain persons. The same objection may be raised against geology by a Protestant Fundamentalist, or against discussion of cancer in a house where somebody is dying of that disease. This is most certainly a problem of tact, decorum and sensibility; but is it a legitimate bar against debate at all times in all places? I doubt it. Those who cannot enter the arena of political debate without being ``hurt" by the first manifestation of conflicting ideas certainly should leave that arena, but they have no real justification in demanding that the arena should be closed down.\\
On a fourth level of intelligence is the anti-free-love \emph{syllogism}. This is based on observation of or experience with various bounders, scoundrels, crumb-bums, etc. --- i.e. exploitative or invasive persons --- who have used (or misused) free love as a rationalization for their predations. The syllogism takes the form:\\
\indent Step 1: X, who is in favor of free love, is a bounder, scoundrel, crumb-bum, etc.\\
\indent Step 2: Y is also in favor of free love.\\
\indent Step 3: Therefore Y is also a bounder, scoundrel, crumb-bum, etc.\\
This is one of the modes of invalid syllogism, as Aristotle knew 2500 years ago; semanticists call it the fallacy of uncritical inference. The weakness is revealed by keeping the form and substituting terms, e.g. Step 1: X likes Jello and is an anti-Semite; Step 2: Y also likes Jello; Step 3: Therefore, Y is also an anti-Semite; ir Step 1: X smokes grass and blows great horn; Step 2: Y also smokes grass; Step 3: Therefore, Y also blows great horn. This is not thinking but revery.\\
On a fifth level, anti-free-love ideology takes the forl of asserting that freedom allows something reprehensible to happen. In the old days, this was supposed to be homosexuality, nowadays, it is supposed to be heterosexuality. Since nobody in anarchist circles would dare to raise the first claim at this time, I will deal only with the second; the two are functionally identical. If we are to exclude heterosexuality on the basis that FADA (Faggot and Dyke Anarchists, New York) or some other Infallible Authority informs us that it is indecent or neurotic or somehow immortal, etc. we have violated the very basis of libertarian thought. Others are then free to demand the suppression of homosexuality on the same grounds; although this sounds incredible today, such anti-Gay bias was once fairly widespread and might be revived at any time. In short, the fact that heterosexuality is currently considered evil or shameful is not in itself grounds for suppression.\\
Free love justifies voluntary, non-coercive homosexuality and voluntary, non-coercive heterosexuality. It cannot justify invasive homosexuality or invasive heterosexuality, e.g. rape or fraudulent seduction. And it cannot justify any attempt to abolish either homosexuality or heterosexuality.\\
A sixth form of anti-free-love ideology takes the form of historical analysis, or, as Karl Popper calls it, \emph{historicism}. This is the anti-evolutionary assertion that the origin of a thing is its ``nature" from which it can never diverge. Thus, if Relativity is of Jewish origin (Einstein being of Hebraic ancestry, although an agnostic), it is therefore ``Jewish physics" and thus untrue; this was a popular line with the Nazis once. Similarly, many anarchists believe that the historical evidence that the State originated in conquest is enough to discredit it; but it is \emph{not}. One must still prove that the State \emph{remains} an exploitative institution (which isn't hard to show, actually).\\
Anti-free-love ideology, similarly, is on unsafe grounds in claiming that free love was invented by men and is therefore a male plot against women. The reply is: 1. Origin is not essence or we would all still be fish. 2. The assertion is factually false; early free-love advocates Victoria Woodhull, Fanny Ward and Emma Goldman were \emph{not} men in drag. 3. Freedom is not a plot against anybody.\\
A seventh form of anti-free-love ideology asserts that sex is not itself important. This may or may not be true (\emph{degustibus non disputandam}), but it is irrelevant anyway. Free love means that a person may choose whatever sexual pattern seems preferable, so long as others are not harmed; this includes anything from homosexuality, heterosexuality, masturbation, triolism, orgies, etc. for those who think sex \emph{is} important, toe celibacy for those who think sex is \emph{not} important. Proclaiming this right is as vital to the celibates as it is to the swingers, if we are ever to have a libertarian society.\\
The eighth protest against free love is based on the claim taht \emph{sexual freedom} (as distinguished from sex itself) is too unimportant to be worthy of debate. While this school of thought does not explicitly reject freedom, it argues that sexual self-determination is merely ``individualistic" or only ``bourgeois" or some way foul and contemptible. This seems to me a total misunderstanding of neurology and psychology.\\
\emph{A person who can be sexually coerced has no identity.} The sexual function is so intimate, so biologically deep and interconnected with kinesic and proxemic self-definition, so ``touchy" and delicate, that when sexual self-determination is surrendered there is little selfhood left to fight any other battle. The woman who can be seduced by any scoundrel who comes down the pike; the devotee of a supernatural religion who has given up sex entirely at the behest of shaman or priest; the rapist with no ability to postpone gratification --- these are all strangely zombie-like types, schizoid and sometimes outright schizophrenic. If you do not ``own" your genitals --- if the Church owns them, or any stronger person in the room owns them, or some ``irresistible impulse" owns them --- you do not possess anything like self-identity.\\
The demand for sexual self-determination is as basic as one's own skin, one's nerves, one's bioplasm. By comparison, the right to own the product of one's labor, or to write and speak without censorship, are less important, because external. Sex goes right to the core, inside the ganglia, down to the cellular level. If ``they" (society, the State, whoever) have you ``by the short hairs," they \emph{really} have you. The battle for self-ownership begins right here; and if one surrenders on this level, there is hardly any spirit left to fight for the externals of freedom. Every tyrant knows this intuitively, and sets taboos on the sexuality of each people he hopes to rule.\\
The ninth argument against free love is based on the allegation that all men are scoundrels. (This is a variation on the syllogism discussed in point four and/or the \emph{argumentum ad hominem} of point two.) There are several answers, as follows:
\begin{enumerate}
\item "Allness" statements are semantically invalid. The lengthy, and somewhat technical, mathematical proof is given in Korzybski's \emph{Science and Sanity}; but, briefly, the arguments offered against anti-Semitism or anti-black prejudice in any liberal high school civics text apply here also.
\item Empirically, the equation \emph{all $m = K(s)$}, all men are in the class of scoundrels, is not accepted by \emph{all} humans (certainly not by men), nor even by \emph{all} female humans, nor even by \emph{all} feminists. The attempt to set policy on the basis of such a claim is, then, an endeavor by a minority to enforce their own phobias on the majority, as in the old puritanism, all forms of institutionalized racism, or tyranny in general.
\item Even if all men were proved to be scoundrels, this is not a \emph{refutation in principle} of free love, but merely a point of prudence or policy. That is, the principle of freedom would still stand but intelligent women would avoid make contact; they would still have the liberty, under anarchist theory, of sexual self-determination in choosing between celibacy, lesbianism, masturbation, bestiality, etc., just as the ``scoundrelly" men would have the liberty of choosing between similar alternatives. The liberty to decide remains a firm anarchist principle, even if wisdom strongly decrees that certain choices shoud be avoided. Cf. heroin of guns.
\end{enumerate}
As a sub-case of argument nine, it may be asserted that under the present patriarchal conditions, any sex between male and female gives all the advantage to the male, encourages him to be fraudulent or dishonest, discriminates against the woman, and generally leads to an ever-closer approximation of making ``all men are scoundrels" become, in fact, true. This ignores the evidence that many persons of both sexes are able to resist the temptation of scoundrelism, however strong that temptation may be. (Cf. pacifists in our jails, Germans killed for helping Jews escape, honest businessmen, politicians who refuse bribes, etc.) But this, again, is a matter of prudence and judgment, not of libertarian principle. Men and women, even under these authoritarian conditions, retain the inalienable right to decide for themselves whether to choose homosexuality, heterosexuality, monosexuality, asexuality or whatever pattern suits them; and this includes the right to decide whom they will and will not trust.\\
There are other arguements against free love but they are explicitly non-anarchist (authoritarian) and do not need refutation in this place at this time. To an anarchist aware of anarchist theory the issue remains clear: Love is either free or coerced. The former represents anarchism and the latter represents tyranny.