\documentclass[12pt, onecolumn, letterpaper, oneside]{book}

\usepackage{authblk}
\usepackage[autostyle]{csquotes}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[sc]{titlesec}

\usepackage[square,numbers,sectionbib]{natbib}
\usepackage{chapterbib}
\usepackage{titlesec}
\usepackage{indentfirst}

\titleformat{\chapter}[display]
  {\rmfamily\scshape\bfseries}{}{0pt}{\Large}
\titleformat{\section}[display]
  {\scshape}{}{0pt}{\large}
  
 \usepackage{scrextend}


\usepackage{suffix}

\newcommand\chapterauthor[1]{\authortoc{#1}\printchapterauthor{#1}}
\WithSuffix\newcommand\chapterauthor*[1]{\printchapterauthor{#1}}

\usepackage{fancyhdr}
 
\pagestyle{fancy}
\renewcommand{\chaptermark}[1]{\markboth{}{\uppercase{\itshape #1}}}
\fancyhf{}
\fancyhead[L]{\rightmark}
\fancyhead[R]{\thepage}
\renewcommand{\headrulewidth}{0pt}
 

\makeatletter
\newcommand{\printchapterauthor}[1]{%
  {\parindent0pt\vspace*{-25pt}%
  \linespread{1.1}\large\scshape#1%
  \par\nobreak\vspace*{35pt}}
  \@afterheading%
}
\newcommand{\authortoc}[1]{%
  \addtocontents{toc}{\vskip-10pt}%
  \addtocontents{toc}{%
    \protect\contentsline{chapter}%
    {\hskip1.3em\mdseries\scshape\protect\scriptsize#1}{}{}}
  \addtocontents{toc}{\vskip5pt}%
}
\makeatother

\title{No Governor}
\author{Vol. I, No. 1.\\Editor Robert Shea}
\date{Spring, 1975}

\fancyfoot[L]{Summer, 1974}
\fancyfoot[R]{No Governor}


%\setlength\parindent{0pt}
    \setlength{\columnsep}{1cm}

\begin{document}
\sloppy

\maketitle

\vspace*{\fill}

\begin{flushright}
\emph{Re-edited using \LaTeX $\,$ for the 'Patafizikhood of Eris Esoteric.\\
If you have any comments or detected any mistake, please contact us at:\\
anarchao--at--discordia--dot--fr.\\}
\end{flushright}

\begin{flushright}
\emph{Thanks to Mike Shea for giving us permission to do this.}
\end{flushright}

\vspace*{\fill}

\tableofcontents

\chapter*{Notes}
\chapterauthor*{Robert Shea}

\blockquote{``There is no governor present anywhere."
\par\begin{flushright} --- Chuang Tzu \end{flushright}
}

After looking and asking around, it appeared to me that no magazine strictly devoted to anarchist ideas and opinion now exists in this part of the world, and that there is a need for one. There are publications devoted to news, to discussions of practical experiences and to polemic, and all of them include some theoretical discussion. But no periodical is devoted just to discussing and developing the underlying ideas that unite --- or divide --- the anarchist movement. We need a fuller understanding of what the term anarchism means. Many who would agree with the sentiment, ``Off the government," have no idea that their own pet proposals might bring back government under a different name. Many are unaware of the problems that will probably arise in anarchist practice, the multiplicity of legitimately anarchistic points of view, the range of possible means to abolish government, the relative merits of the various options.\\
There's a need to express the old ideas that are still valid in new language, to reach contemporary society. There are unresolved questions which need to be further debated. There are new conditions, new opportunities and new dangers to which anarchist ideas must be adapted. To provide an arena in which the ideas and writings of anarchists can meet, be exchanged and clash, I undertake to publish \emph{No Governor}.\\
Many anarchists feel a great impatience with theory, with ideas, with words and discussions. And this is healthy. Words, ideologies, abstractions, generalizations have frequently been used by authorities as instruments of control over subject populations. In a sense, anarchism is an expression of the wordless, vital impulse of rebellion that leaps up irrepressibly in the human heart. Like Zen and Taoism, which are ultimately anarchistic in their implications, anarchism goes beyond words. Yet, anarchism needs words to define itself, to distinguish itself beyond any possibility of confusion from government-oriented political philosophies, to free itself from authoritarian tendencies. It's quite possible to be an unconscious anarchist, but the conscious anarchist has the considerable advantage of knowing what he or she is doing.\\
\emph{No Governor} needs four kinds of help:
\begin{enumerate}
\item It needs to be read. One of the most important responsibilities of an anarchist is self-development as an anarchist. The words on these pages will have life and value only if you let them into your mind. If their only function there is to arouse opposing thought, they will still have served their purpose.
\item It needs comment and criticism. After reading this issue, please write us about the magazine and about specific articles. Letters from readers will, I hope, be a very important part of all future issues.
\item It needs articles. What kind of articles? Well, as we've been saying, the focus in \emph{No Governor} will be on opinion and ideas. Intellectual, theoretical articles. The six pieces in this issue give a pretty good idea of what we will be looking for. And we'd like to publish articles commenting on current events from an anarchist point of view; articles describing practical experience in organization and agitation and drawing strategic conclusions; articles proposing anarchist solutions for some of society's current problems such as war, ecological destruction, crime, poverty, sexism, racism; articles discussing human psychology and sociology from an anarchist standpoint; articles suggesting anarchist strategies for changing society.
\item It needs money to help with the cost of printing and mailing. Anyone who can't pay for a copy of the magazine and who wants one, can get one issue just for the asking. Free sample copies will also be sent to anyone you designate. But there will be no such things as a free subscription. Each issue has to be requested individually, in order to get it free. From those who can pay, I ask 25 cents an issue or \$1 for four issues. Those who contribute articles to \emph{No Governor} will receive subscriptions of indefinite and unforeseeable length.
\end{enumerate}
It has been difficult to organize my time to get this first issue out, so it's labeled ``Summer, 1974" up to page [23] and ``Spring, 1975" thereafter. So this is actually the Spring, 1975 issue. I see reason to think that I am a better organized person now than I was last year, so I hope that the next issue will be Summer, 1975, and that I can follow a quarterly publishing schedule thereafter.\\
So, here is a magazine dedicated to the promotion of anarchist consciousness. My own biases as an anarchist are pacifistic and individualistic, but I will try to publish what I consider good anarchist writing, regardless of whether or not I agree with the position it takes. I take full responsibility for editorial judgments, including all the mistakes.

\include{should_we_cooperate}

\include{the_greedy_gurus}

\include{the_anarchist_movement}

\include{doing_anarchism_yourself}

\include{the_necessity_for_pacifism}

\include{free_love}



\end{document}


















