\chapter{The Greedy Gurus and the False ``New" Left}
\chapterauthor{Tom McNamara}

The American public has a long history of being preyed upon by salvation sellers. If you look back, you'll come across Bible-bangers of the stripe of Cotton Mather, Billy Sunday, Aimee Semple McPherson, Father Divine, Daddy Grace and many other less-known fleecers of the flock. The air-waves still reverberate with the booming voices of the fundamentalists, Oral Roberts (recently become ``respectable" since he has allied with conservative Methodism), Garner Ted Armstrong (who runs a multimillion-dollar church empire that rakes in countless millions a year), conk-haired Reverend Ike (who tells his black followers to get off welfare and ``God" will give them a Cadillac) and Nixon's official evangelist, Billy Graham (whose most recent worry was that his ``God" was going to destroy him with a cataclysm, which gives you an idea about how important he thinks he is).\\
Recently, particularly among the young offspring of America's muddled middle class, it has become stylish to rebel against all this. For about a decade, new left mysticism, fueled by a desire to end the Vietnamese war, was very much in vogue. It was a glamorous way of rebelling and there was a secular religiosity, very much unrecognized, in demonstrating, attending meetings, making flaming speeches about burning issues, writing for movement media, concocting ideologies to fight about with competitive groups. The mass media devoted oceans of space to it all as a sign of its liberalism and everyone involved convinced themselves their commitment to the revolution was heartfelt and lifelong. The disillusionment set in heavy as the 70s dawned and the goals of the pseudo-left revolutionaries suddenly turned into just another flavor of pie in the sky. To counteract the despair and nothingist feelings, to fill the void left by the seeming lack of anything real to commit themselves to, thousands of young people have turned to religious cultism of various unorthodox brands in order to absorb more mysticism and to get the answer to it all.\\
If ``religion is the opiate of the people," as Marx stated, these new synthetic cults are the cheapest, most adulterated kind of street-junk heroin. Their leaders are the most hypocritical charlatans and exploiters and, without fail, fit very snugly into the left cliche ``lackies of the imperialistic ruling class." They are exponents of acquiescence to the oppressive system, worshippers of the status quo, mystagogues of the most crass order, hocus-pocus artists who make palm-readers look like saints by comparison, and are the direct descendants of the carnival ripoff snake oil sellers and other mountebanks. In your modern time they've gussied up their acts with the trappings of scientism, in some instances, using the credibility of gullible millions to weave an electronic mythology of ESP, flying saucers and supposed contacts with spacemen to bilk the ignorant of millions upon millions of hard-earned dollars, creating a very luxurious lifestyle for themselves. Some, like two-ton butterball boy avatar Guru Maharaj Gee-whiz, even have the affrontery to state that since they are ``God" themselves, they deserve to ride in Rolls Royce automobiles and live like kings. They are not only the darlings of the mass media, but of a good part of the old new left as well.\\
By now we've all heard more than enough about the boy ``God" and his Divine Light Mission, which runs up a truly fantastic electric bill to produce the sleazy neon glaze that passes for illumination to the dark minds of his easily-impressed followers. As with the Moslems --- ``There is one God and Mohammed is his prophet" --- there is only one (thank whatever!) Gooroo Gee-whiz and Rennie Davis is his very vocal barker. Davis, star of the Chicago Seven trial and other movement happenings, enhoys an extension of his time in the limelight and his role of apologizer for the Gooroo and his various enterprises. Some people have an insatiable need for power trips and publicity, and the more absurd the proposition, the more challenging to their ability to rationalize their involvement and explain it. Anything as long as they are at or near the center of vast attention.\\
Even when the Gooroo was discovered trying to smuggle watches and illegal bundles of cash into India, Davis and the other mahatmas of this crass and materialistic faith racket were able to laugh it all off in one way or another. True believers will take anything. At least for a while, and most of them are guilt-ridden enough to feel much lighter after they are separated from their valuables by these smiling spiritual thieves and con-men. If this cunning theft were all these string-them-along artists were up to, it would be easy enough to write them off as somewhat intelligent economic criminals. But they are concerned even more with offering soporifics to the hungering minds of their sheeplike followers, blunting their human concerns with changing not only their own spiritual status in a real way bit in taking action to change the horrifying inequities in the real and tangible world around them. The closest the Divine Light Mission comes to social action is re-selling the goods ripped off from new believers in their Salvation-Army-like thrift store. (A thrift store is one in which destitute people are paid little or nothing to sell merchandise usually repaired by other unfortunates who are paid starvation wages, if anything, so that the exploiting organization can reap the profits with little actual overhead.)\\
Although Rennie Davis' promotion is currently the hottest show in town, there is no shortage of hip peddlers of instant enlightenment, meditation --- transcendental and otherwise --- brainwave adjustments, positivity courses, ESP lessons, psychedelic mysticisms, Scientology, UFO cultism, dope theologies, radical psychotherapies, cures and healings and various kinds of blessings, adjustments, both mental and physical, and a thousand and one variations on old and tried (and found worthless) schemes. What few stop to realize is that if the society they live in were not so atrociously rotten, all these vapid panaceas would not only be unnecessary but would be positively laughable. Only the true-believing political groupies, whose spectrum runs from the old and new lefts of various shapes and sizes to the Young Americans for Freedom and beyond even talk about social reform, and that's almost always after the revolution (their equivalent of heaven on earth). These groups have more in common with the pseudo-religions than their members could ever let themselves realize without dropping out. They require true belied, acceptance of almost everything on faith, adulation of all-knowing leaders (Marx, Stalin, William Buckley) and a total acceptance of authoritarianism. This is where anarchism differs radically from all of them. There are no authorities in anarchism and, typically, there is a healthy disrespect for any widely-held opinion.\\
Where the left has fallen down is exactly where the cultists have filled in, gaining converts by the droves. Although both groups put off satisfaction in the now until much later (``after the revolution" or ``in heaven"), the left political groups rigidly changed when the present power structure is overthrown along the explicit correct lines of their particular brand of absolute authoritarian socialism. Some are so blunt about their fanaticism as to sound like the hellish documents of the Nazis. \emph{The Second Page}, a newspaper published by a group in San Francisco, joyously states:
\blockquote{While everybody working towards successful revolution has a responsibility to deal with their personal difficulties as best they can, it is against the interest of the revolution for the revolutionary part to devote much energy to such matters. We don't have time to focus on inner group introspection and ``group-dynamics." We can't afford to concentrate on personal despair. We can't personally resolve the misery of a class fall, or the inequities of our society.}
Hold all the hurt in, folks, don't get blood on the party banner, the revolution will be all over in a hundred years, more or less. No wonder the lemmings desert such slave ships and run for the nearest spiritual snug harbor they can find!