\chapter[The Necessity for Pacifism]{Do an-archists know how to read? or \\The Necessity for Pacifism}
\chapterauthor{Joffre Stewart}
\fancyfoot[L]{Spring, 1975}

\section*{1.}
An-archism begins in non-violence. The logic that tells us that an-archism is rooted in non-violence parallels the chronologic which makes Lao-Tzu the oldest an-archist of record with his Tao Te Ching dated as far back as 2500 years ago. An-archism begins in China but this is not necessarily an argument for cultural diffusion. Many people discover the idea of autonomy for themselves which is almost the only way and then relate to (the sociology of) an-archism when it is brought to their attention.\\
Zeno of Citium, founder of the Stoic school in Greek filosofy, was an an-archist, probably not the first, because he learnt from the Cynics. The filosofical cynicism of the ancient Greeks was apparently a more agitational and social change thing than what is meant by ``cynicism" today. Today it connotes passivity and resignation to things as they are, a wisdom that is no harnessed to doing anything good. Thus E. Howard Hunt cd be a cynic as well as might Lenny Bruce, who, for all I know, carried a draft card, paid taxes, and engaged lawyers: he was not an activist. What these Greeks had to say tended to be negative to slavery \& violence and the civilization that was elaborated on them. You cd not distill a Red Army Fraction from their theories.\\
The next important an-archist in the history of nonviolence (nonviolence seems to have started as an-archism) is Jesus. Mark 10: 42-43 defines an-archism:
\blockquote{$^{42}$ Jesus called them to him, and saith unto them, Ye know that they which are accounted to \emph{rule} over the Gentiles, exercise \emph{lordship} over them; and their great ones exercise \emph{authority} upon them.\\
$^{43}$ But so shall it not be among you:}
As long as an-archists are unable to see that Jesus is one of us, an-archism is in trouble, and, considering the stubbornness of resistance to perceiving this, B I G $\;$ T R O U B L E. Jesus is the most widely disseminated teacher in the world, the most respected (if only superficially) in the Western world, and perhaps second only to Marx in the world as a whole.\\
The failure to dig Jesus as one of us means that we don't know how to read.\\
The failure to appreciate Jesus as one of us means that we underestimate a latent mass potential for an-archy. It means we are throwing away tremendous ``capital." This failure means that we have not penetrated and overcome the Establishment brainwash \& interpretation that ignores, subverts and perverts Jesus' ANTI-State, ANTI-Authority orientation. And this failure to overcome, and therefore to incorporate, Jesus in our history with Bakunin, Joan Baez, John Lennon, Anne Hutchinson, etc., means that we ourselves are not only the victims of this status quo propaganda, but that we cooperate in our victimization. It means that we are suckers for ruling class propaganda. If we are suckers for ruling class propaganda, do we have the strength of mind to prevail in the first place? And if we did gain without assimilating Jesus to our history and comradeship, then do we understand enuf about brainwash, propaganda, and ``co-option" to keep and advantage once gained?\\
The acceptance of Jesus is a test for the survival of the an-archist movement.\\
Now, Jesus, the above indicated Greeks, and Lao-Tzu were an-archists (an-archism \& non-violence being identical) because the polities in which they lived were so violent. An-archism was developed as as solution to this problem, \emph{WAR}, not as a justification for going to war. If those ancients had nonviolent reasons for becoming an-archist, then the radioactivity in the bones of so many of us shd be reason enuf to make us, if not super-pacifists, then at least pacifists, which is the beginning of adequacy in the post-Hiroshima situation. And one is certainly not adequate to an \emph{an-archist} solution if, like Tyrone Walls, you are uncertain how many megatons to deploy after you let the `ruling class' maneuver you into that situation.\\
Notice, that in the above, I use history to b r o a d e n consciousness. The Marxist uses history to \emph{narrow} consciousness, e.g., defining Imperialism in terms of \emph{capitalism} (you'd think Imperialism was born yesterday). This \emph{narrowness} tends to fanaticize consciousness, tends to blind them with righteousness to their own use of Power, and it leaves them fighting among themselves abut capitalist-ANTIcapitalist USSR, China, etc.

\section*{2.}
Since an-archism originates, logically, in the negation of coercion-violence, and chronologically, as a solution to war-slavery, the class analysis is not essential to a viable theory and practice, War makes slaves. Almost every black into ameriKKKa came as a prisoner of war. (1) Class does not establish conquest (2) Conquest establishes class. (3) Class does not establish itself. (4) Economics (labor + `exchange') is not coercion (5) Economics + Authority  = economic coercion (``class") (6) Taxation is economic (7) Taxation is economic coercion (8) Taxation is economic coercion but not ``class" (rent interest profit). Marxists cannot deal with this --- (6), (7) --- because they \emph{WANT} to be guilty of it (9) Law perpetuates conquest (10) Law perpetuates ``class", etc. (If there is any oversimplification in (1)--(10), the truth is in the oversimplification, not in obfuscation.) Marxists in particular are especially ill-trained to understand (1)--(10). Their economic analysis is no more than an excuse to get into politics, therefore an excuse for making war, for gaining State power, and therefore for shooting at each other like Marxist Russia and Marxist China who are into proving once more that \emph{war is in the nature of the State}. But there is no better blinder than Marxist economic analysis to prevent Marxists from seeing that they are demonstrating the an-archist maxim that \emph{war is the nature of the State}. Russian and Chinese Marxists won't even be able to see it in the maximum illumination of mutual exchange of a million megatons. The trouble with this is that we all die for their blindness.\\
It is useless to point out that Marx mentioned primitive accumulation or agreed in part or in essence with (1)--(10) above, because the essence of \emph{Marxism} does not consist in what is reasonable in Marx's economic writings which are irrelevant. \emph{Das Kapital} is the particular economic analysis which is considered the base of the Marxist system. There is nothing in \emph{Das Kapital} that suggests parliamentary procedure, yet this is the most unquestioned, unshakable foundation of Marxist practice. It may be that there is \emph{nothing} in \emph{any} of Marx's writing regarding parliamentary procedure; it may only be assumed in some letters where the task, for example, is to suppress FREEDOM NOW (an-archism). There is nothing in \emph{Das Kapital} that suggests resort to State power, yet Marxists can sooner believe in virgin birth, that not using the State. There is nothing in \emph{Das Kapital} that persuades the reader to accept obligatory servitude, yet Marxists, whom one might at first assume to be libertarian because they are fighting against the Selective Service System, surprisingly, but not infrequently support the principle of conscription themselves. Something in \emph{Das Kapital}, regarding the \emph{corvee}, might suggest that Marx wd be unsympathetic to coercion of labor but Marxists cannot dream of NOT paying Nixon's salary. \emph{Das Kapital} is not prejudiced against pacifism, but Marxists think pacifism ridiculous.\\
Marxism always means what is hostile to FREEDOM NOW. Marxism always means what is hostile to FREEDOM NOW regardless of what may or may not be quoted from the young of old Marx. The political monster, Marxism, ultimately derives from the spirit and practice of Marx, which, as the spirit and practice of Authority, went to war against FREEDOM NOW wherever it crossed his path. In Paris he opposed Proudhon. In Berlin he opposed Stirner. In the context of the First International, he fought Bakunin. And in the conflict with anti-Authority people, Marx rigidified his theory in the more nonhuman, deterministic, nonlibertarian (i.e. ``scientific") directions. The monster became more monstrous as it developed and survived in conflict with anti-Authority forces.\\
(A wishy-washy an-archism cannot deal with this).\\
One readon for this (this outcome) is that we live in the wake and backwash of wars. I grew up, at least in part, in the patriotic backwash of The Big I and then there was The Big II to saturate movies and TV and date the times.We do not date the times by coronations and not really by Jesus since he was absolutely opposed to lordship \& domination, but by orgies of patriotic bloodletting; Post (-World) War I, Post (-World) War II, World War III (pre-Armageddon)... Anti-Establishment authoritarians (and others) of the 19th Century looked to the French Revolution as a model and we live in the backwash of the Russian Revolution which is used to define new terms for domination and spin off new models for violence. And before the new terms of self-deception are established we may see the old ones applied to the new model: the Stalinist system, for example, analogized as ``Bonapartist". With a pacifist consciousness reinforcing out an-archist one (how did they become separate?) we can really dig in and attack smash destroy the assumptions of both the patriotic and the labor-chauvinist models for social violence.\\
Another reason why Marxists are always against FREEDOM NOW may be owing to ethnic origin. Carl Landauer is a socialist, not an an-archist, but in his \emph{European Socialism} he has this to say:
\blockquote{``He [Bakunin] considered the Germans and the Jews the two races with the most authoritarian spirit, and since he regarded this spirit as his deadly enemy, his hatred of the German Jew (i.e. K. Marx), who stood for a rather centralized form of socialism, was very intense, at least in the later phases of their relationship."
\par\begin{flushright} Page 128. University of California Press. Berkeley \& Los Angeles. 1959. \end{flushright}
}
We shd keep in mind that on the official German record, the atheist Marx (author: \emph{A World Without Jews}) was of Lutheran origin, not Jewish, otherwise he cd not have qualified to become a Doktor. Some U.S. Marxist sects may be 80\% Jewish. The eastern European character of U.S. Jewry may mean that the authoritarian practice of Marxism may have been influenced by the factious, discriminating method of theological disputation of that ghetto Jew who replaced the Talmud (or whatever) with the writings of the Herr Doktor professing Karl. Hence, sectarianism as we know it.\\
One cd go on forever both detailing and conjecturing what is wrong with Marxism and Marxists. It is impossible to be too devastating in attacks on Marxism, especially since Marxism orients what is called ``the movement", including the so-called ``anti-war movement". However, attacks done ineptly can be counter-productive. And attacks on Marxism are not attacks on Marxists --- our relations to them shd be pacifistic, of course.

\section*{3.}
If the unity of an-archism and nonviolence (anarcho-pacifism) were as obvious, in the above, as I wd hope it were, then the following questions wd not be asked, but they are asked:
\begin{enumerate}
\item Can people dominated by colonialism get independence by pacifistic means?
\item Isn't it better to make a few gains by violent or parliamentary methods rather than attempt the nonviolence that wd take longer?
\item In time of danger isn't it necessary to defend workers' rights against attacks from the upper class?
\end{enumerate}
1.-2. The first two questions can be answered with one expression: Mahatma Gandhi. The subcontinent did separate from Britain by pacifistic means, means that owe something to the an-archism of Tolstoy and the libertarian individualism of Thoreau. But not enuf, because separation from Britain was achieved by the political form of \emph{independence} rather than in the nonpolitical-ANTIpolitical form of an-archy. The only proper goals for nonviolent action are those consistent with STATELESSNESS (FREEDOM NOW!). It is the grossest perversion of nonviolence to use it to reorganize the forms of domination, and this was evident on the subcontinent long before Gandhi (a liberated person?) exploded the nuclear device that can be intercontinental in 1979. India split from Britain by splitting in 2, and 8 million died in a matter of months, not years like it took to achieve the same gross in WWI. Kashmir War. Two with China. Then Pakistan split in 2 in the bloody affair of Bangla Desh. This is quite counter-productive to the anti-archic aspiration because this aspiration is to \emph{reduce} the number of sovranties on Terra to zero, not to increase them from zero to 3, as in the subcontinental case. Just as an-archism in the unself-contradicting sense, is nonviolent, so nonviolence, to be consistent with itself must be an-archist. Non-violence is not a method of multiplying and enlarging armed forces.\\
Peoples dominated by colonialism \emph{can} get independence by pacifistic means but is the human situation improved now, now that they rape \& slaughter each other by lakhs, millions, crores? Is it worth a Gandhi-made H-BOMB exploding over your head? Is it worth an India-China nuclear war that omnicides the human speecies?\\
The Indian independence struggle, credited, properly, to Gandhian nonviolence (tho there was mutiny in the Indian navy and perhaps incipient guerrilla warfare) took about 27 years and cost 8,000 lives. It is not inconceivable that armed struggle cd have cost 80,000 lives or 800,000. The Indo-Chinese struggle has been on and off for much longer than a generation, at least 800,000 died \emph{during the U.S. intervention alone}, an-anarchists cannot expect security from Marxist sectarianism in North Viet Nam, both Nams have conscription, and Saigon is capital of the largest computerized police State in the world, thanks to the American taxpayer. Before the agreements were signed to start the cease-fire war, it was revealed that the U.S. had plans to stay in Nam 10 more years, and it may well be that the Council on Foreign Relations has no intention ever of abandoning the ameriKKKan presence in Viet Nam. Nonviolence does not necessarily take longer than armed struggle, it costs less in economic drain and in lives, and the gains of exchanging one form of domination for the other are about as illusory in one case as the other.\\
\emph{Parliamentary methods?} Such methods separated ``Nigeria" from Britain, and the war that followed, 7 years later, to reduce the only \emph{legitimate} government in Nigeria, that of Biafra, (was preceded by racist electoralizing) was conducted more intensely than in Viet Nam, and was more frightful, if possible, than the fission that produced Bangla-Desh. Is this what is meant by \emph{``a fes gains by... parliamentary methods"}?\\
Marxists support national liberation movements because Marxists are always against FREEDOM NOW, and the Pearl of the Antilles makes their \emph{venceremos} example. Armed struggle in Cuba was of shorter duration if you count from Moncada, than in the India case --- and the whole world nearly paid for the gains on October 22, 1962. A choice for nonviolent struggle cannot be regarded as as mistake even if nonviolent resistance were still unconcluded in Cuba, about 20 years later. However, the struggle wd be pointless if `making a few gains' left the principle of sovranty intact, or exchanged subordination to the U.S. State Dept. for subordination to the Kremlin, as in the armed struggle case.\\
Peoples dominated by colonialism shd not have to do, proportionately, as much as they do. We are in the belly of the monster. Surely more can be done than throwing Dow Chemical files into the streets, if only more of it. Most of those who term themselves \emph{anti-imperialists} are into ``organizing", not action, into education, too little of which is education for \emph{DIRECT ACTION}, and into making prayer to that which delivers domination instead of wrecking the delivery systems of domination.\\
An-archists are much more lacking than Marxists in what concerns giving specific attention to anti-colonialism, so it is not to be wondered that Marxists get ahead in this area. When an-archists think of a transnational conference, they think of Europe. Not even Latin America is considered. Those an-archists who compromise themselves with passports shd think of having the next transnational congress in India, in the presence of the Sarvodaya movement which is (wishy-washily) an-archic. The Atheist-Who-Walked-with-Gandhi (a liberal) is another possible host for such a subcontinental gathering.\\
\emph{3. Emergency defense of workers' rights?} The best way to answer this is to point out that the upper class (or whatever) is in position to attack workers' rights, today, because workers and others in previous times made themselves content merely to defend those rights \emph{and did it in such a way as to postpone Utopia}. \emph{The State delivers domination} (not Wall Street --- the USSR has no Wall Street) and the pacifist attitude --- equally an an-archist attitude --- of TAX-REFUSAL NOW is always a defense of the workers. The an-archist case is one of prophylaxis: \emph{SMASH THE STATE} before it gets pregnant with a Hitler, Stalin, Franco or whatever. Socialist repression (NSDAP, CPUSSR) seems more efficient than much we've seen in modern times, tho the CIA sis no slouch (Indonesia, Chile). A `successful' defense of workers' rights by Leninist means, may be the worst thing that cd happen to workers.\\
CAVEAT: removing the State as a means of delivering domination may be too narrow an anti-Authority approach if offered as the unique solution to racism-sexism. This kind of caution is, of course, even more relevant to the kind of socialism that overemfasizes economic causation and defers direct action solutions for racism-sexism. That kind of thing, more than anything else, shd be recognized as a crypto-racismsexism.\\
Since an-archism is anti-Authority, it shd be easier for an-archists to have more flexibility than economic determinists and tangle with racism-sexism both as manifestations of Authority and violations of an-archic equality.\\
Nonviolence delivers more means for dealing with these problems than does armed struggle which is counter-productive (or ridiculous) if it can be applied at all.
\begin{center}
F R E E D O M $\;\,$ N O W
\end{center}