\chapter{The Frying Pan}

\emph{Reviews of Libertarian Publications Recently Received\\}

\noindent \textsc{Bayou La Rose.} Reprints useful information and theoretical pieces for an anarchist audience. Issue number 2 includes articles on a nuclear accident in Alabama, the Trilateral Commission, anarchist organizing, the anti-nuke movement in West Germany, health hazards of low-level radiation and I.W.W. history. Sent on request, but a donation of \$5 asked for a 5-year subscription. Copies of articles worth reprinting also requested. Bayou La Rose, Box 58A Rushing Route, Mountain View, Arkansas 72560.\\

\noindent \textsc{Black Flag.} Organ of the Anarchist Black Cross. A lively and comprehensive international anarchist news magazine. Vol. V No. 7 covers the ``Persons Unknown" conspiracy case in Britain, slave labor in the Soviet Union, the Spanish labor situation. Albert Meltzer weighs in with ``The Myth of the Revolutionary Party," in which he assaults the notion that revolutions have to be led by a vanguard elite. Subscriptions are \$14.50 per 12 issues from Black Flag, Over the Water, Sanday, Orkney, KW17 2BL.\\

\noindent \textsc{Black Star.} An Anarchist Review. A publication of the Social Revolutionary Anarchist Federation. Volume 1, No. 6 includes articles on the anarcho-syndicalist trade union federation (CNT) in Spain, the Castro regime, the 1977 Canadian general strike. Subs are \$3 for one year, \$1 extra for airmail. Black Star, box 
92246, Milwaukee, Wisconsin 53202.\\

\noindent \textsc{Cienfuegos Press Anarchist Review}. Number 2. Runs to 64 pages and is packed with a wealth of articles covering a wide spectrum of anarchist interests. Articles include Nagorski's ``History of the Anarchist Movement in Poland"; Kropotkin's essay on ``Anarchism" from the Encyclopedia Britannica; ``B. Traven, Master of the Revolutionary Novel" plus lots of book reviews. \$2 from Cienfuegos Press, Over the Water, Sanday, Orkney, KW17 2BL, U.K.\\

\noindent \textsc{The Dandelion.} A libertarian journal. In the first issue publisher Michael E. Coughlin wrote that he chose the title ``because this little yellow plant reminds me so much of liberty. The dandelion doesn't need any special care or treatment to make it flourish. If we but leave it alone, it will take care of itself." Volume 2, Number 6, Summer, 1978 includes welcome critiques of positions recently taken by Murray Rothbard. Published quarterly, four issues for \$3.50. Michael E. Coughlin, Publisher, 1985 Selby Avenue, St. Paul, Minnesota 55104.\\

\noindent \textsc{The Diagonal Relationship.} The personal journal of Arthur D. Hlavaty. The January, 1979 issue, number 8, includes essays on such topics as the editor's multiple selves, the reissue of \emph{The Principia Discordia}, the semantics of the word ``fuck," professionalism and other conspiracies. Available for \$1, letter of comment, trade, artwork or other plausible excuses from Arthur D. Hlavaty, 250 Coligni Avenue, New Rochelle, New York 10801.\\

\noindent \textsc{Libero International.} Issue number 5, September, 1978, appears almost two years after issue number 4, but the editors promise to continue publication. This issue is devoted to historical studies of Bakunin's visit to Japan in 1861 and about Osugi Sakae, a 20th century Japanese anarchist who revered Bakunin and in many ways resembled him. As always, \emph{Libero International} sheds light on vastly important aspects of the world and its history that are little known in the Occident. No fixed subscription fee, contributions requested (and well worth whatever you send). Libero International c/o LIRA-Nippon SIC, C.P.O. Box 1065, Kobe, Japan 650-91.\\

\noindent \textsc{Newsletter on Intellectual Freedom.} Not an anarchist publication, but a product of the Office for Intellectual Freedom of the American Library Association, this bimonthly newsletter is of interest to anarchists and libertarians because it rounds up all news of all current attempts at censorship in the U.S. Featured in the January, 1979 issue, Volume XXVIII, No. 1, is the defeat of an obscenity law in a South Dakota referendum. The proposed law was modeled after one drafted by the infamous Larry Parrish prosecutor of \emph{Deep Throat}. Indispensable if you want to know what is being done to prevent us from reading and seeing whatever we want. \$8 per year from the American Library Association, 50 East Huron Huron Street, Chicago, Illinois 60611.\\
                                                                                      
\noindent \textsc{Open Road.} News relating to anarchism gathered from all over the world. The only tendency that seems to be neglected is right-wing anarchism or libertarianism. Issue Nine, Spring 1979 covers the anarchist movement in Spain, prison conditions in Canada, childbirth at home, rapes in doctors' offices and an interview with the Living Theatre in Italy. An indispensable publication. No subscription rate, free to anyone who wants it. A donation of an hour's pay at regular intervals or a lump \$50 is requested. The Open Road, Box 6135, Station G, Vancouver; B.C. V6R 4G5 Canada.\\

\noindent \textsc{Self-Management Newsletter.} Vol 2 No. 1. ``Created to promote an ongoing discussion amongst North American anarchist-syndicalists, anarchist-communists, wobblies and libertarian labor radicals about building a libertarian labor movement," Issue includes articles on the UMW coal strike and the 1977 iron ore strike. Available from Resurgence, Box 801, Evanston; Illinois 60204.\\

\noindent \textsc{Southern Libertarian Messenger.} A compendium of many news items of interest to libertarians. January, 1979 issue covers specific ways taxpayers' money is wasted, individual battles with the I.R.S., many amusing and infuriating instances of government stupidity from all over the world, pithy commentaries by editor John T. Harllee. \$3 a year. Box 1245, Florence, South Carolina.\\

\noindent \textsc{Sweet Gherkins From the Dill Pickle Press.} Vol. VIII, \# 7, October 24, 1978 contains another in publisher Edward R. Johnson's series of ``Observations on Fascism," plus quotes from such writers as Rousseau, Thomas a Kempis, Thoreau, Edgar Lee Masters, Upton Sinclair, Li Hung Chang. 10 cents a copy, \$1 for ten issues. Edward R. Johnson, P.O. Box 822, Muscatine, lowa 52761.\\

\noindent \textsc{Srafederation Bulletin.} For Anarchist Agitators, a publication of the Social Revolutionary Anarchist Federation. Consists almost entirely of letters mostly by anarchists, mostly about experiences, practice, theory. Vigorous exchange of ideas. Reading each issue is like attending a bull session with anarchists of all sorts from all over the world. Available for the asking, \$3 a year suggested. SRAFprint, Box 4091, Mountain View, California 94040.\\

\noindent \textsc{The Thunderer.} A broadside sheet announcing world-wide upheavals attendant on the discovery of the body of Jesus Christ. Cienfuegos Press, Over the Water, Sanday, Orkney, KW17 2BL.\\

\noindent A reprinting of \emph{The Principia Discordia}, the Holy Book of the Discordian Movement is available for \$5 from Loompanics Unlimited, Box 264, Mason, Michigan 48854.