\chapter{The Secret Teachings of George Washington}
\chapterauthor{General Strike}

George puffed the reefer and passed it to Tom Jefferson, who was involved in an act of coitus with Discordia, the black ``slave girl" (so thought the unitiatied) (Discordia Jefferson, incidentally enough, was the great-great-great grandmother of Robin Jefferson of modern-day Fucking Communist Conspiracy fame.)\\
Washington spoke these words in reply to Jefferson's contention that the proposed Constitution would lead to tyranny (which had been Franklin's response, at first): ``No. Because of the 13\cent stamp."\\
\textsc{Jeff:} The 13\cent stamp?\\
\textsc{Geo:} Yeah. I give the Constitution 200 years. after that will be the 13\cent stamp.\\
\textsc{Ben:} That's why we have to see that the Federal Government maintains a monopoly on first-class postal business. Dig?\\
\textsc{Geo:} In 1976 --- on New Year's Day --- the Postal Service (which is what it will then be called) issues the 13\cent stamp --- as an omen of the end, as an endorsement of the Indian hemp plant, a commemoration of the 13 colonies who defied the Crown, as a call to revolution, and as an ``inflationary measure of economic necessity" or some similar fiscal responsibility doubletalk --- as if an institution traditionally devoted to losing money needs to worry about inflation!\\
\textsc{Ben:} Right. And you see, the Illuminati will have engineered the most drastic inflationary spiral in the nation's history during '75 solely to enable the Postal Service to credibly raise the price of mailing a first-class letter to 13\cent.\\
\textsc{Geo:} And all first-class mail will thereafter be transported by air, too. Because, in reality, the 13\cent stamp will be a sign --- an omen of the end, an endorsement of the Indian hemp plant, a commemoration of the 13 colonies who defied the crown...\\
\textsc{Discordia:} And a call to Revolution.
\textsc{Ben:} And millions upon millions of these stamps will be on letters which will be in the air over the continent every day.\\
\textsc{Geo:} So Illuminati agents can point out that there are ``signs in the sky."\\
\textsc{Discordia:} That way we prepare the Bible freaks for the change. And not only that. George, tell him about women's liberation angle.\\
``Let me tell that part," said Ben Franklin, expelling a large cloud of smoke (being as he was into women's liberation vicariously as a closet case drag queen, during odd moments between flying kites in lightning storms and printing them \emph{Saturday Evening Post}. ``One of the secrets to which we Rosicrucians are privy is that the number 13 signifies death and resurrection. That is why there were 13 people at the Last Supper. But they were all men, so..."\\
And so they spoke long into the night about secret things, many of which cannot be revealed until the time is ripe.