\chapter{The Ether Vibrates}
\vspace{-1cm}
\emph{Letters from Readers}\\

\noindent \textsc{Chicago, Illinois:}\\

\begin{flushright} 
\begin{tabular}{l p{8cm}}
1931 & Nel forte Braschi di Roma l'anarchico Michele Schirru viene fucilato, perche scoperto mentre attuava un attentato contro Mussolini
\end{tabular}

In the Year of the BOMB --- XXX\\ \end{flushright}

\noindent \emph{no governoR.S.}:
\indent I particularly liked Anton's bitt for ``FREE LOVE... AND ALL THAT". I hope it gets widely copied.... In regard to McNamara's \emph{Greedy Gurus}, the ``Father" is D\emph{e}vine, not Divine.... Regarding ``... The Necessity for Pacifism", there shd be, not a comma, but a period separating \emph{practice} and \emph{War} (p. [27], line [6] from [top]). On p. [30], line [3 from bottom], \emph{thanks} not \emph{`thaks}.' And in the 2nd line of the concluding CAVEAT: \emph{auti-authority}, not \emph{`anti-aughority'}\\
In ``Doing Anarchis Yourself" you seem to come out for the inheritance of acquired characteristics where you write: ``The anarchist movement, little more than a century old, represents a beginning effort by some members of our species to erase that neolithic authoritarian mind-set programming and try to think about human problems in a new way".\\
More important, perhaps, is the nature of the group. I do not regard myself as more important than the group I may be part of. The group is composed of myself and others in interaction. To regard myself as more important might directly of indirectly connote some carelessness toward, or contempt for, others in the group. Not even Stirner, in his Unions of Egoists, wd have so cavalier and attitude toward others with whom he might be in association (altho Stirner incorrectly conceived an inhenrent conflict between the needs of the Ego --- if he meant Egos other than his own --- and the nature of association). A group, particularly an an-archist group, exists in order to satisfy a need for revolution. Such a group shd not exist merely ``to satisfy the needs of individual members". If that were the case, the group wd tend to become a psychotherapy affair: a waste of time. In a group, as in social intercourse generally, one shd seek a balance between the egoistic and altruistic interests of individuals. And in a balanced relationship, the group might have equality with its members.\\
Your theory of the group cd easily lead to Arlene Meyers' criticism of an-archism as ``an anti-social movement".\\
Another thing; I think anti-Authoritarian groups can be as well defined, at least on paper, as Authoritarian groups. Not to be as well defined can be tantamount to not knowing who or what we are or what we are about. SRAF has principles, but I do not think that the schedule of principles makes it less an-archist. Since I made a contribution to those principles, I think SRAF is \emph{more} an-archist because of them. The trouble was that entities like Tyrone, WAP, etc., participated, seemingly as members, when they did not understand or accept (or know of?) the anti-archic waters. I do not reprove the openness by which this happened, but it shd be obvious that if we are more available for penetration by vigorous, eccentric authoritarian elements, then we will have a much harder time getting ourselves together to do anything, as compared with the efficiency of authoritarians to get themselves together to smash us or to compute society like a prison.\\
I am not one of those hostile to all forms of leadership on principle. If I were, I might have opposed your initiative from which the nameless anarchist horde got started. Of course, I oppose all the leadership, out of hand and sight unseen, that needs coercion, hierarchy, or which expresses racism, sexism. But as regards groups such as you and Arlene discuss, the ideal shd be that of a balanced distribution of the functions of leadership. Failing that, the leadership may be useful or valuable as long as it is correct. (The leadership of organizing individuals in limited economic situations is offered or defended as a principle of organization by some aligned with the Warren tendency of individualist an-archism\footnote{This is not more, Bob, than what you do, when as non-governor, you lay down the rules, format and frequency for participating in \emph{No Governor}.}.\\
A. Meyers, discussing the an-archist movement on p. [16], writes: ``our heads are really the first battleground where we initiate the struggle against the state". And in your article, you write: ``One new thing that has to be learned and is rarely fully appreciate is the role of the individual in anarchism." The role of the individual in an-archism is to initiate the struggle against the State, in h/er/is own head, by renouncing citizenship. By defining ourselves to ourselves and to others as STATELESS PERSONS OF THE WORLD, \emph{and shouting it from the housetops}, as it were, we begin to break some new ground against democracy and patriotism. Theories that were well enf adapted to fighting the autocracy of the Czars may be lacking in strategic emfases in situasions where the citizen is h/er/is own worst enemy merely because s/he thinks of self as a citizen, and where the worst despotisms sell themselves to the public as ``democratic". If we can make an-archism clear to ourselves as STATELESS PERSONS OF THE WORLD, then we can make it clear to others. We certainly DO NOT make an-archism clear to the average Jo(e) if s/he thinks we are somebody `just like h/im/er'. There shd be an important and essential difference between us and this average Jo(e) and just letting Jo(e) know that we do not belong to h/is/er nuclear club (The State) can have more propaganda value than all the an-archist lit printed since 1965. And some of the solutions we are seeking may flow from our own self-concept, once we put it bluntly, relevantly and correctly to our personal selves an to others.\\
Those who renounce citizenship may not be able to carry a passport to an international an-archist congress. (As it is now those go who have the money or transportation.) Until an-archists, as a movement, can get together on it, there is no alternative but for the conscientious an-archist to declare for Statelessness and disqualify Self for the paper that might deliver h/er/im to an an-archist congress abroad. But if, say, 63 percent of the known an-archists sort themselves out by lottery into those who renounce citizenship and those who might travel on the State's paper, there wd be some available to travel that way. As new people renewed or enlarged the ranks of those organized for STATELESSNESS, their lot wd be cast, periodically, with the moiety that did not luck-out of citizenship: they wd have the same chance of lucking out as all those who did not luck-out of citizenship (formally) in the previous lottery. A renunciation of citizenship, is of course, irreversible (how can an an-archist become, let alone re-become, a citizen?) and the pool of an-archists who have not formally renounced citizenship is continually giving up members to explicit STATELESSNESS. This scheme is not meant to provide a loophole to an-archists who have not explicitly renounced citizenship so that they may think they are justified in voting against GoldwaterWallaceReaganApeshit or voting for WAPLibertarianPartyCitizenshit. If if has the faintest possibility of being interpreted that way, then it shd be junked, discarded without any further consideration. The idea is not to weaken an-archism but to strengthen it. If the scheme can be exploited as to support political participation, then the infra-structure of an-archy is too compromised and subverted by Democracy and Liberalism. It wd be better for all who understand an-archism correctly to unreservedly declare for STATELESSNESS and forego any tactical advantage that the lottery might afford the international movement.\\
At any rate, a given lottery pool shd come to term at the end of 9 or 14 years which means that everybody in the pool wd declare for STATELESSNESS and a new 9 yr. pool started with newcomers. (Better, terms of different lengths --- 9 or 14 --- cd be assigned to joiners. This wd ensure a continuity).\\
It is too much like a disgrace that International An-archist congresses are helf with people who arrived on passports. If these Congresses concluded with most or all of those on passport burning them and declaring themselves STATELESS, then the Congresses might serve as a useful demonstration of anti-patriotism, anti-nationalism, an-archism \& human solidarity. I do not say that we shd get rid of such Congresses, questionable as they may be. This is not an attack on the an-archist movement: we need all that we have. But I do say that International An-archist Congresses shd be called where participants from abroad do NOT have tourist card, visa, permit, passport or other politically certified travel papers. Whether such an International Congress cd reach a `quorum' (whatever that might mean for an anti-parliamentary gathering) remains to be explored. But an-archist activity, including ``going limp" when made \& held captive, at check points or border crossings wd practically be assured. A chance to do something about the main impediment to human solidarity. With an an-archist interpretation that \emph{ATTACKS} the worldcitizen-worldgovernement terminology.\\
And there is no reason why we shd not call such International Congresses in places like Gulai-Polye, Changsha, or Chicago (an-archists have beend forbidden by law to enter USA since 1903). We might have as much success getting together without passports in those places as anywhere else.\\
It goes without saying that anarchists, like Jesus, shd be born bastards. The an-archist who has a child who is not bastard has something to explain --- mainly to the child. But the point here is that there shd be a way for a child to be born without automatically  being tainted with citizenship and nationalism: the democratic imposition of identity. Is the immaculate conception possible? At any rate we shd try, and publish the places and ways a child might at least come into the world STATELESS. If the ungrateful brat wants to rebel against its parents and kiss the feet of Grace Kelley, live among the bogey men of Haiti or wallow in the despotism of Israel or USSR, it will, unfortunately, probably have such chance before we can turn the world upside down and make such options available.\\
Arlene Meyers, in her ``The Anarchist Movement --- Dead of Alive?" suggests that a ``lack of direction" contributed to the demise of a Chicago an-archist group. I wd suggest also that that group's failure to meet a challenge of patriotism in regard to a 5th of July parade (Evanston, ILL), also contributed to the group's demise --- internal disintegration --- because a dishonest maneuver was resorted to in order to blunt part of the attack that might have been made on patriotism. The dishonest maneuver was fount out, and the expose, limited as it was, unglued some relationships.\\
In other words, we trick ourselves when we treat patriotism flippantly, condescendingly, lightly, when we treat it as not as important as, say, wage struggles. It is much more serious than wages \& hours struggles except when strikes occur in war industry. At that point they become the same problem, because the workers shd not be trying to improve their wages \& hours position, they shd be trying to destroy it utterly and irrevocably insofar as it is connected to war work. And this workers' attack on military employments shd be delivered with attacks on patriotism. Patriotism is the most dangerous religion around.\\
Near the top of page [13], Arlene speaks of ``integrating our politics? with our daily lives". Since an-archists, as ANTI-political animals, do NOT have `politics', it follows that it is our ANTI-POLITICS that we shd be intergrating with our daily lives.\\
Jim Bumpas, in 3 places in his ``Should We Cooperate?" refers to the ``corporate establishment" which dominates our culture and society. I think ``corporate establishment" is a misleading piece of new left rhetoric, because it is the people who lend money, especially those who lend money to Governments who orient the status quo and draw its basic outlines. In this lite, ``corporta establishment" is more like a cover-up for the WASP-Jewish money what runs it than anything else.\\
Bumpas says: ``It is to our advantage to tip the status quo out of balance". And he says that, by so doing, ``we will not be swallowed up whole as the Bolsheviks swallowed anarchists in Russia" even while the CP in Portugal is swallowing up everything. We shd recognize that the present balance of the status quo probably protects us more that it threatens us, both from ``Left"-authoritarians and the Rightracistreactionaries. The present balance of the status quo is probably the best protection going for most Marxists, because the sect that gets the Power will probably smash all the rest, much like they are trying to dow, and are doing in Portugal.\\
We shd not be trying to ``tip the balance" of the status quo. We shd be trying to collapse it utterly: no authoritarian arrangement for either the political left or the political right to work thru.\\
This is not an argument against cooperation since I probably do more of it (with \emph{Marxists}) than most an-anrchists who read this. The advantage of cooperation is that it gives \emph{Marxists} a chance to disabuse themselves of their stereotypes (frequently, stereotypes indistinguishable from the bourgeois stereotypes). Those of us who have been \emph{Marxists} know what that means. And those of us who understand the difficulties with black-white stereotyping may understand the size of this job.

\par\begin{flushright}
\begin{tabular}{l}
Joffre Stewart\\advocate of the \emph{ANTI}-Christ\\6114 S. May Street\\Chicago, Illinois 60621 
\end{tabular} 
\end{flushright}