\chapter[Anarchism and Magick]{Anarchism and Magick in the Light of Interstellar Neurogenetics}
\chapterauthor{Robert Anton Wilson}

\blockquote{Just as every natural science and every branch of engineering has been revised, expanded and re-defined in the light of atomic and nuclear physics, so it is necessary to expand and re-define every branch of human knowledge in the light of the relativity and fission-fusion capacities of the nervous system... Science, philosophy and human affairs in general operate with inefficiency and dangerous short-sightedness because of the ignorance about the nature of the instrument used, i.e., the nervous system.
\par\begin{flushright} --- Timothy Leary, \emph{Periodic Table of Energy} \end{flushright}
}
This essay assumes, without restatement or argument, the basic factual-theoretical Universe of modern science. Matter as interference-patterns (knots) in energy. Evolution, local and cosmic. Relativity of space-time-matter. DNA as the building-plan of neurogenetic evolution.\\
We assume also the Arrhenius-Crick theory of cosmic panspermia: DNA is distributed throughout the galaxy via nucleotide templates, certainly on meteors, possibly on other celestial objects. (The May 14, 1964 meteor in Orguel, France, contained such a template.) We accept Sir Francis Crick's hypothesis that this panspermia is consciously guided by Higher Intelligences. We prefer Dr. Timothy Leary's interpretation of the motives of such Intelligences (love, curiosity, generalized parental instinct) to Crick's own interpretation (scientific zeal, pride, imperial power-politics).\footnote{Leary evidently models his Higher Intelligences on the erotic-exuberant gods of paganism; Crick, on the jealous Gods of monotheistic Judaism, Christianity, Islam.}\\
When the first DNA arrived on Earth four billion years ago, it began evolving, adapting, adjusting to the gravitational, meteorological and other characteristics of this planet. If Carl Sagan of Cornell is right, there approximately 1,000,000,000 (one thousand million; one billion) planets in our local galaxy on which similar evolution of the DNA has taken place or is taking place.\\
According to Dr. Leary's hypothesis, the evolution of neurogenetic systems follows similar laws on all planets, following the four basic stages of:
\begin{enumerate}
\item ---bio-neural survival-intelligence;
\item ---muscular-glandular emotional intelligence, i.e. rage-fear and similar territorial and domination rituals;
\item ---symbolic intelligence, or logical-abstract ``thinking" (sub-vocal speech);
\item ---domestic intelligence; sexual-amative impulses generalized successively to care for the family, the tribe, the planetary community.
\end{enumerate}
Bio-survival first circuit intelligence is characteristic of unicellular organisms in the amoeboid family; emotional second circuit intelligence, of the amphibians, reptiles, mammals, et al; symbolic third circuit intelligence, of humans, dolphins, whales; domestic fourth circuit intelligence of pack-bonding or pair-bonding species, especially predators who need cooperative intelligence in the hunt.\\
All of these stages survive in contemporary humanity, especially (but not exclusively) in children. In general, the neo-nate knows only bio-survival forward-backward vectors; the toddling infant becomes aware of muscular-glandular energies and finds a vertical position in the up-down of emotional politics, first in the family, then in society; the symbolic ``computer" intelligence begins with childish handling-questioning, and gets programmed during school years; the erotic-domestic intelligence is imprinted after puberty.\\
We find among adults (including anarchists) a surviving preference for one of these modes. Some are first circuit bio-survival types, floaters, drifters, slow to anger, slower to think, usually (these days) tranquilized by pot. Others are second-circuit people, emotional-territorial, investing energy in protecting ideological ``turf," maneuvering up-down relations in political-passionate groupings, perpetually operating the rage-fear glandular reflexes. A few (damned few) are capable of the sustained use of third circuit symbolic intelligence characteristic of the philosopher, logician or scientist. And some are what astrologers call Aquarians, feeding all energy into eroto-domestic fourth circuit harmony in the family, the group, the band, ultimately the planetary-ecological Whole.\\
As suggested in Dr. Leary's \emph{Neurologic}, \emph{Interstellar Neurogenetics}, \emph{Terra II}, and \emph{Periodic Table of Energy}, it is thinkable, possible and in some ways probable that evolution is extra-planetary as well as planetary. In that case, there are higher forms of intelligence, beyond the four already existing. Indeed, we find traces of these higher neurological functions in such deviant groups as shamans, magicians, mystics, Sufis, Rosicrucians (historical, not the San Jose con game), some Buddhist or Hindu sages, some mathematicians and scientists who have pushed themselves beyond symbolic intellect, a few really disciplined and mature LSD-users, etc. Dr. Leary suggests that these higher modes of intelligence include:
\begin{enumerate}
\setcounter{enumi}{4}
\item ---neurosomatic intelligence: the ``wisdom of the body," the somatic Turn On, the rapture or Satori state; \emph{un}conditioned reflexes;
\item ---neurophysical intelligence, to be achieved fully in group-linkages, human-computer linkages, cybords, etc.;
\item ---neurogenetic intelligence, direct perception of the DNA code within;
\item ---metaphysiological intelligence, moving from experimental transfers outside the body (astral projection) to final union with Galactic Mind.
\end{enumerate}
If many planets are still evolving through the first four circuits, many others have probably evolved into the latter four. This is especially likely after the more advanced species of each planet achieve fusion-energy, blast off from the home planet and begin to mingle with more advanced races of the galaxy.\\
Dr. Leary suggests further that all the higher types of intelligence, appearing only sporadically on earth, will consistently appear in consecutive (evolutionary) order among our descendants living in free-fall conditions of zero gravity, extended life, sane decentralization,\footnote{The Second Level emotional-hierarchical reflexes are inappropriate and maladaptive for Deep Space. They are based on gravitational senses within the nervous system. Attempts to transfer the Earthian hierarchical-militaristic mentality to the Galactic level, as in the dismal Star Trek fantasy, are foredoomed. The Bully, Generalissimo, or Politician type, beginning with the first territorial rituals of Permian reptiles 500,000,000 years ago, will become extinct in the next century, although presumably several NASA-style star-ships sponsored by Earthian governments will fail before this principle is understood.} etc. In short, we and all other inhabited planets are pre-programmed, DNA-coded, to evolve all the way from amoeboid bio-survival to Galactic Consciousness.\\
The techniques of magick and mysticism fine down to variations on the principle: turn off the usual neurological circuits; wait; \emph{futique} vision comes. That is, by deliberately suppressing emotional circuits (\emph{niyama}), symbolical-intellect circuits (\emph{dharana}), bio-survival motility (\emph{asana}) and eros (\emph{yama}), the yogi forces his nervous system to mutate. The higher, still-evolving circuits are (at least temporarily) turned on. (This answers the question asked in Dr. Leary's famous 1968 hypno-tape, ``What Do You Turn On, When You Turn On?") The traditional shamanistic approach uses sex and/or drugs (the Amerindian shamans had 150 psychedelics in their larder, according to Dr. LeBarre's \emph{Ghost Dance}; Aleister Crowley, the latest magus of the Western tradition, got by on hashish, cocaine and peyote --- assisted by ritual and somatic training). This is less wearying than yoga, gets results faster, but is much trickier and easily degenerates or relapses into the First Circuit bio-survival floating. (This is ``the Hedonistic Trap" in Dr. John Lilly's \emph{Programming and Metaprogramming the Human Biocomputer}, the ``repetitious silk-lined womb" warned of in Dr. Leary's \emph{Neurologic}. Most hippies fall into it happily.)\\
The great neuro-programmers of the Occult tradition often share the anarchist vision, frequently in a Stirnerite form. (``All is permitted" --- Hassan i Sabbah. ``Do what thou wilt" --- Crowley.) This is not as destructive as it sounds; the externalized, socially-conditioned, absolute moral codes of terrestrial humanity are quite pointless from the viewpoint of the higher intellectual functions.\\
\emph{The relativity of neurology is as basic as the space-time relativity of Einstein}. The olfactory universe if the dog contains thousands of smelly signals not registered by the primarily visual human. The bio-survival organism of the neo-nate neither scans nor transmits the emotional signals so important to three-year-olds (of all ages). The sexy signals (sperm-egg fusion invitations) so overwhelming to the adolescent are literally invisible to the pre-adolescent. Etc. Similary, \emph{the moral codes of the first four circuits become utterly irrelevant on the fifth and higher circuits}, just as terrestrial territoriality is a null-class concept to the Starship voyager.\\
There is no ``morality" on the bio-survival first circuit; the ``morality" of the emotional second circuit is subjective, whimsical, chaotically relativistic. (Ordinary ``morality" in most tribes is emotional politics: Second-Circuit fear-rage spasms: taboo.) Ethics, in the civilized sense of a Confucius or a Marcus Aurelius Antoninus, a Bentham or a Joseph Fletcher, balances Third-Circuit rationality with Fourth-Circuit domesticity. Rational ethics is, in short, a game of symbolic calculation (Spinoza and Bentham were on the right track in trying to mathematicize it.)\\
When the Fifth Circuit opens to neurosomatic rapture, ethics becomes as irrelevant as the Second-Circuit emotional spasms of the child (or the childish adult) are to the Third-Circuit reasoner. Ethics always balances off ``Mine" and ``Thine"; its basic question is always, ``Who has the right to what? Who owns (controls) this?" It assumes the isolation of each nervous system from the space-time continuum. On the Fifth Circuit, the nervous system and its external signals become a seamless unity. ``I, Thou, It are one." \emph{Benevolence replaces ethics}.\\
On the Sixth and higher circuits, the individual point-of-view increasingly disappears, amalgamates, synthesizes with the cosmic point-of-view.

\section*{II}
Magick, then, is the \emph{futique} or ``feedforward" use of the nervous system. The shaman, witch, magus, etc. attunes his/her nervous system to rarely-used, still-evolving circuits. Sex-magick (Tantra, Sufism, Templarism, Crowleyanity) forms a five-circuit two-person resonator, in the manner of an electrical capacitor, for concentrating, storing, magnifying erotic-domestic energies usually dissipated in ``normal" civilized sexuality of the ``wham-bam-thank-you-ma'am" variety. The ``normal" orgasm --- ``the sneeze in the loins," as Lawrence scornfully called it; Shakespeare, more mournfully, dubbed it ``the momentary trick" --- is explosive. Tantric orgasm is undulatory. The former loses energy; the latter stores it.\\
This is metaphysiological, not metaphysical. The word ``metaphysical" describes a null-class. Neurological fission and fusion, like atomic fission and fusion, releases dramatic and unexpected energies, but does not ``transcend" the energy-continuum in any sense. Metaphysics is bad logic, not beyond-logic.\\
Although it sounds spooky to say it, there is a real neurological link between the space-explorations of the 1960s--'70s and the Hedonic revolution also occurring. The sex-drugs experimentation (marijuana-sex, LSD-sex, cocaine-sex, together with the upsurge of interest in Tantra, sensuous massage, vibrators, porno, gimmickry, swapping, etc.) represents an invasion of the Fifth Circuit by millions of non-magicians, non-shamans, non-mystics. The DNA is programming us for the drop-out of free-fall, the transcendence of the up-down, good-evil, strong-weak dualisms of the first four circuits. The majority of astronauts come back neurologically changed --- inclined toward mysticism \emph{or else} suddenly neurotic, baffled, alcoholic, overwhelmed by new neural energies they won't and can't use --- because gravity-less functioning is Fifth Circuit functioning. As long as NASA sends straight, square, WASPy, conventional types into space, this mind-blowing will occur. \emph{Space is for acid-heads, magi, witches, Tantrists. Those entirely programmed by the first four circuits belong planetside.}\\
``You have to go out of your mind before you can come to your senses," said Fritz Perls. The rapture turn-on of the Fifth Circuit, sensory-somatic intelligence, is the drop-out from the first four circuits of conditioned intelligences. It is folly and dangerous to send socially-conditioned WASPs into space. Giving Acid to Nixon is the only equivalent misapplication of technology.\\
The Jefferson Starship has popularized the notion of stealing a starship and escaping Earth. This is like the usual Yippie objection to anarchism --- "Without a government, where will we get our Welfare checks?" The assumption is that ``we" wait for the grown-ups to create something, then ``we" rip it off. Bio-neural, First Circuit intelligence, infantile parasitism. Dr. Leary has suggested, instead, that we build our own starship.\\
In the late '60s, Leary and various others --- including Lord Omar of the Discordian Society --- were investing energy in the idea of oceanic libertarian communes. Fleets of merry anarchists living in international waters. The thought was that the land-mass had been seized by existing governments and only the sea is still free. Defects in this scheme led first to the concept of underwater (submarine) communes and, then, abandonment of the whole plan. \emph{There is no escape on Earth}. Second-Circuit child-adults in government offices have it all under their territorial-emotional domination.\\
Neurogenetic evolution shows that species-dominants and \emph{futique}-mutants cannot occupy the same territory. Earth will probably be unmanageable \emph{without} scientific totalitarianism by 2020, if not sooner. The engineering-technological Utopia, Brave New World, bee-give. Escape from this perpetual Fourth-Circuit ``togetherness" is only possible in space.\\
Every form of social experimentation is now illegal on Earth; methods of control and surveillance are improving yearly, monthly, even daily. All the long-hairs will be barbered; all the revolutionaries jailed in the Gulag or California archipelagos! All the free, wild animals made extinct; all the soft earth covered with concrete; all the swingers led back to Calvinist or Stalinist domesticity.\\
The drop-out from Europe to America in the 16th--19th Centuries was the inevitable consequence of the rise of the national state. The mutant must escape of perish. The territorial-authoritarian Alpha-male cannot tolerate one single free man or woman in the same space-continuum. The drop-out from Earth to Space is equally inevitable. Starseed, Inc., a non-profit corporation for space exploration outside government auspices, has been incorporated in California and is seeking to lease an island from Japan to begin building an escape ship. Rogue scientists, drop-out engineers, artistic-literary folk who can raise money by sales of media rights, will increasingly join Starseed to make the plan feasible.\\
This has probably been the history, and will be the history, on every viable planet. The DNA program is \emph{first}, adapt to the local planet, survive, increase intelligence, master the first four circuits; and \emph{second}, blast off, escape the planet, join the galactic community, master the second four circuits. If we are between the fourth and fifth circuits now, probably half the viable planets are below us, groping upward from amoebahood, and half are above us, evolving toward metaphysiological union with Galactic and, next, Cosmic Mind.\\
A starship of 5000 experimentally inclined WoMen is the last place where neurological (and therefore social, economic, political) freedom can thrive. It is the only place where intelligence can continue to grow, in accord with evolutionary DNA-programming. The meek shall inherit the Earth --- and turn it into a concrete ant-hill. The wild, free, unpredictable, spermatozooic, exuberant and still-evolving must join the Galactic gypsies to survive.\\
As Dr. Leary's mate, the beautiful Joanna, has said, ``Become \emph{very} intelligent. That's our means of propulsion."
 