\documentclass[12pt, onecolumn, letterpaper, oneside]{book}

\usepackage{authblk}
\usepackage[autostyle]{csquotes}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[sc]{titlesec}

\usepackage[square,numbers,sectionbib]{natbib}
\usepackage{chapterbib}
\usepackage{titlesec}
\usepackage{indentfirst}

\titleformat{\chapter}[display]
  {\rmfamily\scshape\bfseries}{}{0pt}{\Large}
\titleformat{\section}[display]
  {\scshape}{}{0pt}{\large}
  
 \usepackage{scrextend}
 \interfootnotelinepenalty=10000

\usepackage{fancyvrb}
\usepackage{suffix}

\newcommand\chapterauthor[1]{\authortoc{#1}\printchapterauthor{#1}}
\WithSuffix\newcommand\chapterauthor*[1]{\printchapterauthor{#1}}

\usepackage{fancyhdr}
 
\pagestyle{fancy}
\renewcommand{\chaptermark}[1]{\markboth{}{\uppercase{\itshape #1}}}
\fancyhf{}
\fancyhead[L]{\rightmark}
\fancyhead[R]{\thepage}
\renewcommand{\headrulewidth}{0pt}
 

\makeatletter
\newcommand{\printchapterauthor}[1]{%
  {\parindent0pt\vspace*{-25pt}%
  \linespread{1.1}\large\scshape#1%
  \par\nobreak\vspace*{35pt}}
  \@afterheading%
}
\newcommand{\authortoc}[1]{%
  \addtocontents{toc}{\vskip-10pt}%
  \addtocontents{toc}{%
    \protect\contentsline{chapter}%
    {\hskip1.3em\mdseries\scshape\protect\scriptsize#1}{}{}}
  \addtocontents{toc}{\vskip5pt}%
}
\makeatother

\title{No Governor\\
		\large \textit{a journal of anarchist ideas}\\
		\vspace{2\baselineskip}
		\normalsize ``There is no governor present anywhere.'' --- Chuang Tzu
		}
\author{Vol. I, No. 2.\\Editor Robert Shea}
\date{Fall, 1975}

\fancyfoot[L]{Fall, 1975}
\fancyfoot[R]{No Governor}

%\setlength\parindent{0pt}
    \setlength{\columnsep}{1cm}

\begin{document}
\sloppy

\maketitle

\tableofcontents

\chapter[Anarchism and Magick]{Anarchism and Magick in the Light of Interstellar Neurogenetics}
\chapterauthor{Robert Anton Wilson}

\blockquote{Just as every natural science and every branch of engineering has been revised, expanded and re-defined in the light of atomic and nuclear physics, so it is necessary to expand and re-define every branch of human knowledge in the light of the relativity and fission-fusion capacities of the nervous system... Science, philosophy and human affairs in general operate with inefficiency and dangerous short-sightedness because of the ignorance about the nature of the instrument used, i.e., the nervous system.
\par\begin{flushright} --- Timothy Leary, \emph{Periodic Table of Energy} \end{flushright}
}
This essay assumes, without restatement or argument, the basic factual-theoretical Universe of modern science. Matter as interference-patterns (knots) in energy. Evolution, local and cosmic. Relativity of space-time-matter. DNA as the building-plan of neurogenetic evolution.\\
We assume also the Arrhenius-Crick theory of cosmic panspermia: DNA is distributed throughout the galaxy via nucleotide templates, certainly on meteors, possibly on other celestial objects. (The May 14, 1964 meteor in Orguel, France, contained such a template.) We accept Sir Francis Crick's hypothesis that this panspermia is consciously guided by Higher Intelligences. We prefer Dr. Timothy Leary's interpretation of the motives of such Intelligences (love, curiosity, generalized parental instinct) to Crick's own interpretation (scientific zeal, pride, imperial power-politics).\footnote{Leary evidently models his Higher Intelligences on the erotic-exuberant gods of paganism; Crick, on the jealous Gods of monotheistic Judaism, Christianity, Islam.}\\
When the first DNA arrived on Earth four billion years ago, it began evolving, adapting, adjusting to the gravitational, meteorological and other characteristics of this planet. If Carl Sagan of Cornell is right, there approximately 1,000,000,000 (one thousand million; one billion) planets in our local galaxy on which similar evolution of the DNA has taken place or is taking place.\\
According to Dr. Leary's hypothesis, the evolution of neurogenetic systems follows similar laws on all planets, following the four basic stages of:
\begin{enumerate}
\item ---bio-neural survival-intelligence;
\item ---muscular-glandular emotional intelligence, i.e. rage-fear and similar territorial and domination rituals;
\item ---symbolic intelligence, or logical-abstract ``thinking" (sub-vocal speech);
\item ---domestic intelligence; sexual-amative impulses generalized successively to care for the family, the tribe, the planetary community.
\end{enumerate}
Bio-survival first circuit intelligence is characteristic of unicellular organisms in the amoeboid family; emotional second circuit intelligence, of the amphibians, reptiles, mammals, et al; symbolic third circuit intelligence, of humans, dolphins, whales; domestic fourth circuit intelligence of pack-bonding or pair-bonding species, especially predators who need cooperative intelligence in the hunt.\\
All of these stages survive in contemporary humanity, especially (but not exclusively) in children. In general, the neo-nate knows only bio-survival forward-backward vectors; the toddling infant becomes aware of muscular-glandular energies and finds a vertical position in the up-down of emotional politics, first in the family, then in society; the symbolic ``computer" intelligence begins with childish handling-questioning, and gets programmed during school years; the erotic-domestic intelligence is imprinted after puberty.\\
We find among adults (including anarchists) a surviving preference for one of these modes. Some are first circuit bio-survival types, floaters, drifters, slow to anger, slower to think, usually (these days) tranquilized by pot. Others are second-circuit people, emotional-territorial, investing energy in protecting ideological ``turf," maneuvering up-down relations in political-passionate groupings, perpetually operating the rage-fear glandular reflexes. A few (damned few) are capable of the sustained use of third circuit symbolic intelligence characteristic of the philosopher, logician or scientist. And some are what astrologers call Aquarians, feeding all energy into eroto-domestic fourth circuit harmony in the family, the group, the band, ultimately the planetary-ecological Whole.\\
As suggested in Dr. Leary's \emph{Neurologic}, \emph{Interstellar Neurogenetics}, \emph{Terra II}, and \emph{Periodic Table of Energy}, it is thinkable, possible and in some ways probable that evolution is extra-planetary as well as planetary. In that case, there are higher forms of intelligence, beyond the four already existing. Indeed, we find traces of these higher neurological functions in such deviant groups as shamans, magicians, mystics, Sufis, Rosicrucians (historical, not the San Jose con game), some Buddhist or Hindu sages, some mathematicians and scientists who have pushed themselves beyond symbolic intellect, a few really disciplined and mature LSD-users, etc. Dr. Leary suggests that these higher modes of intelligence include:
\begin{enumerate}
\setcounter{enumi}{4}
\item ---neurosomatic intelligence: the ``wisdom of the body," the somatic Turn On, the rapture or Satori state; \emph{un}conditioned reflexes;
\item ---neurophysical intelligence, to be achieved fully in group-linkages, human-computer linkages, cybords, etc.;
\item ---neurogenetic intelligence, direct perception of the DNA code within;
\item ---metaphysiological intelligence, moving from experimental transfers outside the body (astral projection) to final union with Galactic Mind.
\end{enumerate}
If many planets are still evolving through the first four circuits, many others have probably evolved into the latter four. This is especially likely after the more advanced species of each planet achieve fusion-energy, blast off from the home planet and begin to mingle with more advanced races of the galaxy.\\
Dr. Leary suggests further that all the higher types of intelligence, appearing only sporadically on earth, will consistently appear in consecutive (evolutionary) order among our descendants living in free-fall conditions of zero gravity, extended life, sane decentralization,\footnote{The Second Level emotional-hierarchical reflexes are inappropriate and maladaptive for Deep Space. They are based on gravitational senses within the nervous system. Attempts to transfer the Earthian hierarchical-militaristic mentality to the Galactic level, as in the dismal Star Trek fantasy, are foredoomed. The Bully, Generalissimo, or Politician type, beginning with the first territorial rituals of Permian reptiles 500,000,000 years ago, will become extinct in the next century, although presumably several NASA-style star-ships sponsored by Earthian governments will fail before this principle is understood.} etc. In short, we and all other inhabited planets are pre-programmed, DNA-coded, to evolve all the way from amoeboid bio-survival to Galactic Consciousness.\\
The techniques of magick and mysticism fine down to variations on the principle: turn off the usual neurological circuits; wait; \emph{futique} vision comes. That is, by deliberately suppressing emotional circuits (\emph{niyama}), symbolical-intellect circuits (\emph{dharana}), bio-survival motility (\emph{asana}) and eros (\emph{yama}), the yogi forces his nervous system to mutate. The higher, still-evolving circuits are (at least temporarily) turned on. (This answers the question asked in Dr. Leary's famous 1968 hypno-tape, ``What Do You Turn On, When You Turn On?") The traditional shamanistic approach uses sex and/or drugs (the Amerindian shamans had 150 psychedelics in their larder, according to Dr. LeBarre's \emph{Ghost Dance}; Aleister Crowley, the latest magus of the Western tradition, got by on hashish, cocaine and peyote --- assisted by ritual and somatic training). This is less wearying than yoga, gets results faster, but is much trickier and easily degenerates or relapses into the First Circuit bio-survival floating. (This is ``the Hedonistic Trap" in Dr. John Lilly's \emph{Programming and Metaprogramming the Human Biocomputer}, the ``repetitious silk-lined womb" warned of in Dr. Leary's \emph{Neurologic}. Most hippies fall into it happily.)\\
The great neuro-programmers of the Occult tradition often share the anarchist vision, frequently in a Stirnerite form. (``All is permitted" --- Hassan i Sabbah. ``Do what thou wilt" --- Crowley.) This is not as destructive as it sounds; the externalized, socially-conditioned, absolute moral codes of terrestrial humanity are quite pointless from the viewpoint of the higher intellectual functions.\\
\emph{The relativity of neurology is as basic as the space-time relativity of Einstein}. The olfactory universe if the dog contains thousands of smelly signals not registered by the primarily visual human. The bio-survival organism of the neo-nate neither scans nor transmits the emotional signals so important to three-year-olds (of all ages). The sexy signals (sperm-egg fusion invitations) so overwhelming to the adolescent are literally invisible to the pre-adolescent. Etc. Similary, \emph{the moral codes of the first four circuits become utterly irrelevant on the fifth and higher circuits}, just as terrestrial territoriality is a null-class concept to the Starship voyager.\\
There is no ``morality" on the bio-survival first circuit; the ``morality" of the emotional second circuit is subjective, whimsical, chaotically relativistic. (Ordinary ``morality" in most tribes is emotional politics: Second-Circuit fear-rage spasms: taboo.) Ethics, in the civilized sense of a Confucius or a Marcus Aurelius Antoninus, a Bentham or a Joseph Fletcher, balances Third-Circuit rationality with Fourth-Circuit domesticity. Rational ethics is, in short, a game of symbolic calculation (Spinoza and Bentham were on the right track in trying to mathematicize it.)\\
When the Fifth Circuit opens to neurosomatic rapture, ethics becomes as irrelevant as the Second-Circuit emotional spasms of the child (or the childish adult) are to the Third-Circuit reasoner. Ethics always balances off ``Mine" and ``Thine"; its basic question is always, ``Who has the right to what? Who owns (controls) this?" It assumes the isolation of each nervous system from the space-time continuum. On the Fifth Circuit, the nervous system and its external signals become a seamless unity. ``I, Thou, It are one." \emph{Benevolence replaces ethics}.\\
On the Sixth and higher circuits, the individual point-of-view increasingly disappears, amalgamates, synthesizes with the cosmic point-of-view.

\section*{II}
Magick, then, is the \emph{futique} or ``feedforward" use of the nervous system. The shaman, witch, magus, etc. attunes his/her nervous system to rarely-used, still-evolving circuits. Sex-magick (Tantra, Sufism, Templarism, Crowleyanity) forms a five-circuit two-person resonator, in the manner of an electrical capacitor, for concentrating, storing, magnifying erotic-domestic energies usually dissipated in ``normal" civilized sexuality of the ``wham-bam-thank-you-ma'am" variety. The ``normal" orgasm --- ``the sneeze in the loins," as Lawrence scornfully called it; Shakespeare, more mournfully, dubbed it ``the momentary trick" --- is explosive. Tantric orgasm is undulatory. The former loses energy; the latter stores it.\\
This is metaphysiological, not metaphysical. The word ``metaphysical" describes a null-class. Neurological fission and fusion, like atomic fission and fusion, releases dramatic and unexpected energies, but does not ``transcend" the energy-continuum in any sense. Metaphysics is bad logic, not beyond-logic.\\
Although it sounds spooky to say it, there is a real neurological link between the space-explorations of the 1960s--'70s and the Hedonic revolution also occurring. The sex-drugs experimentation (marijuana-sex, LSD-sex, cocaine-sex, together with the upsurge of interest in Tantra, sensuous massage, vibrators, porno, gimmickry, swapping, etc.) represents an invasion of the Fifth Circuit by millions of non-magicians, non-shamans, non-mystics. The DNA is programming us for the drop-out of free-fall, the transcendence of the up-down, good-evil, strong-weak dualisms of the first four circuits. The majority of astronauts come back neurologically changed --- inclined toward mysticism \emph{or else} suddenly neurotic, baffled, alcoholic, overwhelmed by new neural energies they won't and can't use --- because gravity-less functioning is Fifth Circuit functioning. As long as NASA sends straight, square, WASPy, conventional types into space, this mind-blowing will occur. \emph{Space is for acid-heads, magi, witches, Tantrists. Those entirely programmed by the first four circuits belong planetside.}\\
``You have to go out of your mind before you can come to your senses," said Fritz Perls. The rapture turn-on of the Fifth Circuit, sensory-somatic intelligence, is the drop-out from the first four circuits of conditioned intelligences. It is folly and dangerous to send socially-conditioned WASPs into space. Giving Acid to Nixon is the only equivalent misapplication of technology.\\
The Jefferson Starship has popularized the notion of stealing a starship and escaping Earth. This is like the usual Yippie objection to anarchism --- "Without a government, where will we get our Welfare checks?" The assumption is that ``we" wait for the grown-ups to create something, then ``we" rip it off. Bio-neural, First Circuit intelligence, infantile parasitism. Dr. Leary has suggested, instead, that we build our own starship.\\
In the late '60s, Leary and various others --- including Lord Omar of the Discordian Society --- were investing energy in the idea of oceanic libertarian communes. Fleets of merry anarchists living in international waters. The thought was that the land-mass had been seized by existing governments and only the sea is still free. Defects in this scheme led first to the concept of underwater (submarine) communes and, then, abandonment of the whole plan. \emph{There is no escape on Earth}. Second-Circuit child-adults in government offices have it all under their territorial-emotional domination.\\
Neurogenetic evolution shows that species-dominants and \emph{futique}-mutants cannot occupy the same territory. Earth will probably be unmanageable \emph{without} scientific totalitarianism by 2020, if not sooner. The engineering-technological Utopia, Brave New World, bee-give. Escape from this perpetual Fourth-Circuit ``togetherness" is only possible in space.\\
Every form of social experimentation is now illegal on Earth; methods of control and surveillance are improving yearly, monthly, even daily. All the long-hairs will be barbered; all the revolutionaries jailed in the Gulag or California archipelagos! All the free, wild animals made extinct; all the soft earth covered with concrete; all the swingers led back to Calvinist or Stalinist domesticity.\\
The drop-out from Europe to America in the 16th--19th Centuries was the inevitable consequence of the rise of the national state. The mutant must escape of perish. The territorial-authoritarian Alpha-male cannot tolerate one single free man or woman in the same space-continuum. The drop-out from Earth to Space is equally inevitable. Starseed, Inc., a non-profit corporation for space exploration outside government auspices, has been incorporated in California and is seeking to lease an island from Japan to begin building an escape ship. Rogue scientists, drop-out engineers, artistic-literary folk who can raise money by sales of media rights, will increasingly join Starseed to make the plan feasible.\\
This has probably been the history, and will be the history, on every viable planet. The DNA program is \emph{first}, adapt to the local planet, survive, increase intelligence, master the first four circuits; and \emph{second}, blast off, escape the planet, join the galactic community, master the second four circuits. If we are between the fourth and fifth circuits now, probably half the viable planets are below us, groping upward from amoebahood, and half are above us, evolving toward metaphysiological union with Galactic and, next, Cosmic Mind.\\
A starship of 5000 experimentally inclined WoMen is the last place where neurological (and therefore social, economic, political) freedom can thrive. It is the only place where intelligence can continue to grow, in accord with evolutionary DNA-programming. The meek shall inherit the Earth --- and turn it into a concrete ant-hill. The wild, free, unpredictable, spermatozooic, exuberant and still-evolving must join the Galactic gypsies to survive.\\
As Dr. Leary's mate, the beautiful Joanna, has said, ``Become \emph{very} intelligent. That's our means of propulsion."
 

\chapter{The Arrogance of Authority}
\chapterauthor{Jim Bumpas}

Does the title of this article make you think I am going to discuss ``authoritarians?" Or the snotty arrogance assumed by many Marxists once they have read a little Lenin or Mao? That is a bit trite for an anarchist theoretical discussion. I want to help develop some ideas upon the distortions produced in our own practice by what I call the ``arrogance of authority."\\
I believe it is a mistake for anarchists to ignore the problem of authority and the related problem of leadership within our own ranks. Likewise I believe it is a mistake to try to obscure the problem as Spanish comrades did in the 1930s by using euphemistic terms such as ``influential militant" instead of ``leader." I do not oppose the substitution of such euphemisms as long as we develop a clear analysis around their usage and as long as we do not allow them to cloud our practice with destructive distortions and to make clarity of analysis very difficult.\\
These distortion in the practice of our Spanish comrades have recently been analyzed in Vernon Richards' ``Lessons," and to a lesser degree one is made aware of some of them in Sam Dolgoff's book on the Spanish Collectives. The arrogance which I want to analyze here is not the arrogance which anarchists assume when they occupy the traditional seats of power in govt and other hierarchical bureaucracies. I do not believe anyone reading this will argue that an anarchist can get a bulldozer to fly if only an authoritarian can be removed from the driver's set and an anarchist installed.\\
Rather, I want to discuss the arrogance that an organization and its members allow to develop which permit any ``influential militant" or ``leader" to climb into any driver's seat and yet retain any former close ties to the anarchist movement. I think this arrogance is very similar to that arrogance exhibited by the Marxist who insists his understanding of Marxism-Leninism-Mao-Tse-tung thought provides the only possible theoretical underpinning of revolutionary action. If an anarchist organization has not already prevent this problem by clearly analyzing its elements, then perhaps ruthless criticism and even ridicule are the best weapons to use against those who would be leaders OVER the anarchist movement.\\
What about leaders? Are leaders destructive and incompatible with our perspectives? And leadership? Should we avoid ``influencing" others? If so, what about propaganda? Education? Exemplary action? Is the only alternative to leadership, and to being an ``avant garde" simply to withdraw into our anarchist communities and try to live consistently with our perspectives? Anarchist islands in an authoritarian sea? There are some rip-off artists ready to promise just such an escape. And there are thousands of people living in such communities who are gaining experience in collective, solidaric living.\\
But anarchism is a set of perspectives which aim towards social change. That is a fact. Therefore, both people and social conditions of a material nature must be altered, changed, transformed, revolutionized. And people must become convinced of the necessity and desirability of such change, whether by experience, logic, insight or other means. The social action required to spread this experience, logic and insight throughout society involves the following process which I describe as ``leadership":
\begin{enumerate}
\item the development of class analysis;
\item the clear articulation of the destructive effects of class society;
\item the identification of both structural and human symptoms of these destructive effects; and
\item the selection of relevant and effective projects which aid us in any of the above 3 steps, or which accomplish direct social change in themselves.
\end{enumerate}
Obviously, if everyone in society followed these steps, or similar steps, revolutionary leadership would not be necessary. Or any leadership for that matter. For revolutionary leadership to be most effective, revolutionary organization is required. In spite of what I have just said, I believe we can and must avoid any situation where we refer anyone to our ``leader" (singular or plural) when asked, ``Take me to your leader." Our organizational ideal will best conform to our perspectives when we put into practice the old Wobbly slogan, "We are all leaders." If we have achieved this in our own organizations, then we might be able to develop the experience and perspectives we need to achieve the same in society as a whole.\\
This is really the crux of the differentiation between anarchists and our organizations as opposed to authoritarian, hierarchical organizations. Anything which develops the initiative, self-reliance and self-confidence of anarchists and of people in general reduces the present popular dependence upon authority. The tendency of our activity is and should be to make this dependence disappear. We are not like a Marxist-Leninist group trying to shift dependencies from ``them" to ``us" so that we may step up into the driver's seat of that bulldozer. We do not want to develop authorities on anarchism, or among anarchists.\\
Nevertheless, many anarchists appear to exhibit behavior which is conditioned by the authoritarian dependencies and which serve to perpetuate them, albeit laced with anarchist rhetoric. I refer the reader to the Bulletin of the Social-Revolutionary Anarchist Federation for both the evidence of anarchists exhibiting authoritarian behavior and manner of thinking as well as for the efforts of anarchists to combat the effects of authoritarian conditioning and to help one another develop out of the conditioning into more free modes.\\
Some anarchists assert that one is either an anarchist or one isn't. They assert that theoretical discussion, patient explanation and mutual exchange of ideas are both unnecessary and wasteful bullshit. This attitude is not limited only to those who want to lead us right to the barricades for a fight to the death against the force of reaction. It is also expressed by those who want to lead us into exemplary action of other types, such as developing alternate communities or media action for its shock value. Or maybe just propaganda is the object. In any case, this attitude exemplifies an incredible arrogance.\\
On the one hand, they appear to say that an anarchist appears on earth by some mystical process in a perfect form that discussion and theoretical perspectives cannot improve. The only task of anarchists is to take ``the word" to the people and internal education of ourselves is wasted activity when there are so many people out there, thirsting for our message. This arrogance assumes that anarchists need no further education once we decide that the word ``anarchism" best describes our social perspectives, and that we can educate people to accept something closer to our vision of a free society even if we cannot educate them to be anarchists. This is very like the Christian authoritarian evangelist who ``knows Christ" and you cannot understand unless you also know ``HIM".\\
Other anarchists who exhibit a similar arrogance believe that clear development of our perspectives and the words we use to communicate our perspectives are unimportant. They have internalized the Madison Avenue dictum that words can be made to mean what we want them to mean. In the resulting confusion, they believe people will pick up anarchism off the shelves as a result of hyping the product name and associating it with any and every popular concept which has permeated our popular culture. To concretize this analysis, I need only list the words ``hippie," ``free market," ``Woodstock Nation," ``anti-terrorist," etc. Anarchism is a product to be sold on the market. So what if everyone is confused about its actual meaning, or the consequences of anarchist practice in society? Some of these ``anarchists" are more interested in, say, the ``Free Market" than they are in anarchism and are just playing both sides of the street.\\
The arrogance of this form of anarchist authoritarianism postulates an ignorant public, well manicured for these salesmen of anarchy by the social conditioning of the present culture. Why not take advantage of these facts and manipulate the public ourselves, just like everyone else does? Too bad for the manipulators, but people are smarter that these types give them credit for being.\\
A third type of anarchist authoritarian is more sophisticated than the other two. These types seldom participate in discussions in the SRAF bulletin, so a reference to that document may be unavailing. This type is really a synthesis of the other two, and an intellectual improvement. If the first two types of authoritarian anarchists were the ``immaculate conception" type and the ``sales hype" type, this third type might be called the ``embarrassed intellectual."\\
This type does not deny the value of internal education, but insists that discussion must be ``deep" or ``significant," and above all, ``correct." They are embarrassed by any evidence that anarchists are not perfect. (My god! People might even associate ME with these other people who call themselves anarchists!) They are embarrassed by the honest attempts of people (less capable than they) who are struggling with their anarchist perspectives and their efforts to articulate these perspectives. They express the opinion in many ways that if anarchists must describe their perspectives in public (even to an all-anarchist public) that their discussion should not contain any embarrassing errors. Such errors should be edited or censored from any publications so that no one can discover that maybe all anarchists have not developed their perspectives to a consistency equivalent to what Kropotkin achieved. I used to think Marxists were the only ones afraid to take a position on any question until they were sure they had the correct position. Anyway, insistence upon perfection before one acts or opens one's mouth certainly is not leadership, because the definition of perfection is socially determined as a result of the interchange among people and with the material conditions the face. These embarrassed intellectuals know, of course, what is and is not significant and worthy of putting forth to the public The inference is that we should defer to their judgment in theoretical matters.\\
Aside from whatever information appears in the SRAF bulletin, perhaps all anarchists cannot benefit from the discussions of our perspectives there. I refer to that bulletin in my article here because it is the only place I can think of where one can taste the expressions of all varieties of persons who are anarchists on this continent. Men and women (though mostly men), the very young, the very old, professional students and other academic intellectuals, semiliterate, craftspeople, blue and white collar workers, unemployed, professional or middle-income people, and welfare recipients. All these people treat each other as equals in dignity and respect and try to develop the experience to help each other build the means to create a free society.\\
I am convinced our own internal education is fully as important as whatever education benefit we may contribute to society as a whole. No matter how embarrassed one might be by the expression of, and attempts to articulate, our perspective by others, the process is very important for us all. Anarchists are not perfect. We have no infallible leaders to follow. We need to develop ourselves and our perspectives as much as anyone in this society We are as much affected by social conditioning as the most contented bourgeois. The difference is, we recognize a need for change and are struggling for the means.\\
Anyone who tries to prevent this process of education or to assert that it is unimportant, or worse, is really acting in an arrogant, authoritarian manner. Such an assertion is really an attempt to coerce those of us who are struggling with our perspectives to defer to the authority of the intellectuals, or some other ``standard-maker." I have benefited greatly from the mutual process which has been taking place in the SRAF bulletin, as well as from similar types of discussions in face-to-face meetings. Even though I recognize the bulletin is not agitational literature in itself appropriate to hand to non-anarchists on the street, the bulletin is also a tool and a source for ideas we can use to develop our own leadership potential so that we may provoke, influence, and entice others to develop their own initiative and self-reliance in society as a whole.\\
This is the kind of leadership we need to develop. If we do develop it, the ``leaders" will cease to exist. I hope by this process we can all become more fully aware of the authoritarian distortions produced in us by our conditioning. Any arrogance or self-righteousness about our own perspectives has got to be a product of this destructive conditioning. I hope we cease putting each other down for our inability to express ourselves in the best way possible. Our organizational efforts should be designed to provide us with solidarity and mutual support in our struggles --- both internal struggles with our education and perspectives inside our organizations, and external with the forces of reaction and confusion. Libertad o muerte.


\chapter{The First Anarchist (poem)}
\chapterauthor{Kerry Thornley}

\begin{verbatim}
out on a limb
out on a limb
with jerusalem slim
     jerusalem slim
lest ye be perfect
gentle and wise
passing no judgements
on those other guys

yet scorning the hypocrite
and the letter of law
and clearing the temple
like you never saw

down with the high priest
out with the prigs
and woe to ye vipers
ye capitalist pigs

the time is now coming
the time is now near
the time has not ever
been elsewhere than here



and no greater love has any than this
to lay down your life in return for a kiss
to live it for others and die it the same
to cast off your bushel and lift up your flame

nailed to a limb
nailed to a limb
o jerusalem slim
  jerusalem slim
you are the brother
who launched the great struggle
with words they now twist
and verses they juggle
\end{verbatim}

\chapter{The Simple Life}
\chapterauthor{Joffre Stewart}

\section*{Prefatory note}
Most (practically all) of the material below was delivered in a class on the Simple Life, taught by one Bill Ray at Francis W. Parker School (Chicago), on TUE 20 MAY 75. Out of 12 students, 2 helped to complete poem at end by putting match to wee flag. The student publication aired some debate regarding the burning of the Constitution, etc.

\section*{1}
Presenting an-archism from the angle of the Simple Life, as I do here, will have 2 limitations. It will not develop the an-archist critique of Marxism. Marxists-in-power have been the most dangerous enemies of anti-statism, of an-archism, since 1917. The 2nd limitation is that little will be said of the world-wife an-archist movement as it exists today.

\section*{2}
Some an-archists, called \emph{primitive an-archists}, have turned to anthropology and turned their attention toward certain nonliterate societies as the \emph{natural} societies that best take human nature into account. They want to pattern their an-archy on the best of such societies.\\
Partly because of this, the Simple Life may be seen as an an-archistic criticism of civilization. Or, a criticism of civilization, leveled at it by the an-archist spirit. It does not matter what kind of civilization, all it has to be is a civilization. This is because too many things that a wise person in the simple life can do without, are things collected by foul means, by car, exploitation, slavery, coercion. Therefore it happens, quite logically, that the simple life appears as an expression of pacifist an-archism. An' sho 'nuff: the first an-archists who left a record were pacifist an-archists: Lao-Tzu, the Greek Cynics \& Stoics, and Jesus of Nazareth. So the first an-archist, anarcho-pacifist, was a Chinaperson, Lao-Tzu, who, after expressing his way of life in the poetry called the Tao Te Ching, is said to have quit his job as a government burocrat and went off to the mountains to live the simplest life.\\
The Cynics, and the Stoic, Zeno of Citium who learned from the Cynics, tended to come from the lowest class in Greek society that had enuf liberty \& leisure to filosofize. That is, they were a cut above the slaves, barely out of slavery. The point is that their filosofy was a rational expression of their class interest because they were against slavery. They understood that to be against slavery you had to be against was because slaves are prisoners of war. \emph{Peace \& Freedom} are one and the same thing. The person who tries to sell you one without the other is a politician and the politician is the greatest danger to both.\\
The Cynics are renowned for their simplicity in a civilization that was pretty uncomplicated as it was compared to the Sears Towers, Apollo space ships, TV, computer spying, long forms, legal jargon, traffic jams, wiretaps, ice cream cones and radiation diseases of the present civilization. But on of these Cynics lived in a tub. When one of them, perhaps the same one, saw a boy drink water with his hands, he threw away his cup. When Alexander the Great asked the wise man what cd he do for him, the wise man asked the Great Man to step aside so he cd get some sun.\\
Jesus was nonviolent, and because he understood his position, he was not a liberal, he was an-archist, he was a pacifist revolutionary. The Son of Man, as Jesus called himself --- Jesus practically never called himself `Christ' and there is an an-archist reason for this --- lived such a simple life that he had nowhere to lay his head. He told the rich people to give it all away and follow him. He said that it wd be easier for a camel to go thru the eye of a needle than for a rich man to enter the kingdom of heaven. By ``Kingdom of Heaven" Jesus meant an-archy, paradise regained. Tho Jesus worked as a carpenter, he contradicts the Protestant ethic where he asks you to dig on the lilies of the field: they neither spin nor sew, and yet, blooming naked, they got it all: Solomon is all his glory cd not compete. Tho we know less about the sex life of Jesus that we know about god's --- or rather because of that --- it is not necessary to interpret the Nazarene version of the simple life the way the priesthood has. Jesus indicated that he came to encourage abundant living and abundant living can mean all of sexual fulfillment.

\section*{3}
The facts about these people in the classical tradition are usually taught, if at all, in a superficial way that ignores and covers-up the fact that they were against politics, Governments, wars, and bosses. But a lot of schoolteaching time lay be spent on slave owners like Aristotle or totalitarians like Plato. Jesus belongs to classical times but Jesus is not taught because he is controversial, but however, controversial for the wrong reasons, the wrong reasons being like those unimportant reasons that separate Protestant from Catholic. Jesus is no controversial for the right reasons: the an-archist teachings which caused the Protestants \& Catholics to gang up on the people who \emph{did} correctly understand Jesus and destroy them in 1535. Those people were called Anabaptists. And to this day, the survivors of those 16th Century Anabaptists, called Amish, Brethren, Mennonites, Quakers, may prefer the horse to farm machinery, will not address persons by titles, will not send children to secondary school and in other ways show a preference for the simple life, whether you agree with the details or not.\\
Christianity, its divinity schools \& complicated theologies exist in order to suppress the important truths about Jesus; and they exist because Jesus' anti-Statism is suppressed. To know and live the simple truths of Jesus is to destroy Christian civilization.

\section*{4}
The an-archists of modern times who are attracted to simplification as part of the solution also bring nonviolence with them as part of the solution.\\
A graduate of Harvard named Henry David Thoreau, who wrote an \emph{Essay on Resistance to Civil Government}, does not fit exactly into either the an-archist or pacifist category. But he was an anti-slavery man and \emph{Freedom is what an-archism is all about.} Freedom is what an-archism is all about, so as a LIBERTARIAN ANTI-SLAVERY INDIVIDUALIST, Thoreau was very close to an-archism, and by practicing this \emph{near}-anarchism, he has taught the modern world as much about behaving non-violently as has Jesus. And, as with the Cynics, Stoics, and other ancients, this convergence of nonviolence and anti-Statism leads into the simple life.\\
For on thing, the simple life was necessary for Thoreau because he did not want to pay taxes. TAXATION IS ROBBERY. The person who taxpays regularly is in slavery. In order to not pay taxes, Thoreau lived on the margins of the economy in such ways that the tax-collector cd not collect.\\
And I live where Thoreau lived: on the margins of the economy. I have not made a payroll since August 1943, therefore, I have not rendered payroll tax, \emph{income tax}, since 1943. I did not pay for the Korean war, the Viet Nam war, and I ma not paying to keep those people in jail who resisted the Viet Nam war. I did not pay to kill the Rosenbergs. I did not pay the Bay of Pigs invasion, and I did not pay the CIA to kill Malcolm X, JFK, J. Edgar Hoover, and 1,000,000 Indonesians in 1965. I am not paying Ford-Kissinger to send Marines to commit violations in Cambodia. I did not buy the 2 atomic BOMBS that Nixon wanted to give the French to dump on the Vietnamese. I did not pay Nixon's salary. The taxpayer may be the biggest criminal you can find, and you shd consider it an honor to be persecuted for o p e n l y refusing to pay taxes if ever it gets around to that.\\
The Russian Count, Leo Tolstoy, reckoned that you don't fight against war by going to war, and he recognized The State as a (very necessary) cause of war and so became an an-archist altho he did not wear the label. And like an-archist pacifists since Lao-Tzu 2500 years ago, his beliefs took him into the practice of the simple life. He renounced all the fashionable ideas of aristocratic, ruling class life, and began to live the simple life of the peasants on his estates, and began to do constructive things that helped out this simple life.\\
Tolstoy recognized that he did not belong to the power arrangements that arrange for masses of people to kill each other in war, and because he was not part of those power arrangements, he was therefore not a citizen of Russia of of any other country. He was a STATELESS PERSON OF THE WORLD, much life the Stoic, Zeno of Citium, thousands of years before him. It might also be said that Thoreau was not a citizen of the United States or of any other country, but Thoreau had a more sly way of saying it which was self-contradictory as well.\\
I renounced citizenship around Nagasaki Day 1950, declaring myself A STATELESS PERSON OF THE WORLD. The point is that the pacifist an-archists, inclined toward the simple life, tend to have a better idea of who they are in regard to the claims of The State on their identity. The an-archists who are coming from and who are closer to a labor and socialist background are less clear about who they are and this can be a weakness when it comes to refusing to be drafted. The proletarian class concept tends to deal with a part of a person's humanity rather than all of it. This concept has seen the grossest failure in 2 world wars. Even Kropotkin forgot that he was anti-capitalist and took sides in WWI. It is easier for anyone who is not a pacifist to slide into any kind of war that for someone who is pacifist. In this day and age, the an-archism which is not uncompromisingly pacifist is especially weak and irrelevant.\\
In this regard, it is impossible to overemfasize the desirability of attacking and destroying patriotism. One of the biggest failures o an-archism today is that it is not shooting down patriotism. Patriotism is more dangerous than racism, altho the racism that drove so many Jews into Zionism may destroy the world. On the other hand, one way to attack racism is by attacking patriotism, and I will try to show what I mean later...

\section*{5}
Right now I will deal with the Equal Rights Amendment. The Equal Rights Amendment will have women thrown in jail for not carrying draft cards and more than double the amount of cannon fodder available to the Pentagon. I don't believe in jails, I don't believe in war, and an-archism is against all the Law that organizes jail \& war. So what does that mean for the Constitution?
\begin{center}
(\emph{burned copy of Constitution})
\end{center}
By the light of this flaming Constitution, you see plainly why an-archists can never think of going for the Equal Rights Amendment. If there is no Constitution, there is no place to put an amendment. Liberating society from the State liberates women from a male dominated institution that inherently puts women at a disadvantage. An-archism is feminism's best bet. And without the State, there is nothing to be patriotic about. It is as simple as that.

\section*{6}
This brings us to present time. The 2 an-archists best known to Americans are Joan Baez and John Lennon. This is an ambiguous distinction because they are better known as singers than as an-archists. And being so well known interferes with their an-archism because they are rich and this is too fancy for some an-archists from the labor, libertarian socialist background: they tend to discriminate against John \& Joan on account of their ``class." I think class discrimination is unfortunate no matter who does it to whom. Rather than ignore John \& Joan as an-archists, as some IWW types do, I think it is better to acknowledge them as an-archists, and relate to them by criticizing their shortcomings. That way everybody benefits.\\
The people criticize Joan Baez for her money, yet have a lot to learn from her: she pulls down the flag and cuts it to pieces with her tongue. She gives patriotism the trouncing it deserves. She obstructed an induction center with her body. And her money helped reprint some an-archist classics.\\
But she is insincere and incomplete when it comes to tax-refusal: she makes a gesture of refusal but leaves money in the bank where the government can seize it. This is what she shd be criticized for, not merely for having the money. If the pacifist Joan Baez were for real as a pacifist tax-refuser, she would find it very difficult to get or keep money and her life style might be more simple than I think it is because she might not have a choice. My understanding is that her husband left her because her life style reflected her high earnings...\\
I don't know the calculation on John Lennon's personal appointments: the image I have is that they are modest even tho he's rich. What is worse, however, is that he is not getting in trouble for conscientiously refusing taxes.\\
But he and the other Beatles, for an-archist reasons, did reject the Queen's honors. This means a lot more than when Viet Nam vets throw their medals back at the Government. He is not Sir John Lennon and he cd be. The Beatles also tried to destroy fashion, but they were trapped in their leadership and only succeeded in setting new fashions. There is a difference between being a fan/a follower and being an an-archist.\\
But John Lennon at least acknowledges in more ways than one that the simple life is the direction in which he ought to go. Not only does he tell us this in his song writing, but he associates the simple life with out-and-out atheist an-archism. The title of they lyrics is \emph{Imagine}: 
\begin{center}
\begin{tabular}{l l l l}
Imagine there's no heaven & & &  imagine no possessions\\
it's easy if you try & & & I wonder if you can\\
no hell below us & & &  no need for greed of hunger\\
above us only sky & & &  $\;\;$ a brotherhood of man\\
imagine all the people & & &  imagine all the people\\
living for today... & & & sharing all the world...\\
\\
imagine there's no countries & & & you may say I'm a dreamer\\
it isn't hard to do & & & but I'm not the only one\\
nothing to kill or die for & & & I hope someday you'll join us\\
and no religion too & & & and the world will be as one\\
imagine all the people\\
living life in peace...
\end{tabular}
\end{center}
John Lennon is up for deportation because he is such a good poet. AmeriKKKa does not want the best and tries to get rid of it in any way convenient. But the an-archist movement is weirdly lacking in solidarity whenever it cannot rally behind such a case as his.

\section*{7}
The Peacemaker movement has developed the notion of simple living into one of its active principles. The Catholic Worker movement lays even more stress on what it calls \emph{voluntary poverty}. Both movements are an-archist, both are pacifist, and relations between them are cooperative and mutually reinforcing. In terms of the circulation of their periodicals (\emph{The Peachemaker}; \emph{The Catholic Worker}) these two account form by far the bulk of the an-archist movement in ameriKKKa (but this way of reckoning can be tricky...) In terms of lasting power and continuity, these 2 \emph{are} the an-archist movement in ameriKKKa. But in terms of theoretical discussion, their activity is minimal, essentially unhistorical, and unsatisfying for people who have wide-ranging needs to know. What they do do is the propaganda of the deed: the nonviolent ACT.\\
The Peacemaker movement began around 1948. It came into existence to break draft laws, especially to encourage \emph{you} not to register for the draft. Nor do its members do war work. By 1952, the nonpayment of taxes was as firmly a part of Peacemakers as refusal to register for the draft. From the beginning, Peacemakers adopted the notion of inner personality transformation and of nonviolent revolution but the idea of nonviolent revolution did not clearly take on an-archist characteristics until sometime between 1962 and 1972. THE IDEA OF SIMPLE LIVING MUST HAVE BEEN THERE ALL THE TIME, but the recognition of it as one of their characteristic properties and practices came slowly and built up gradually in time. A column on Simple Living often appears in the Peacemaker publication.\\
The Peacemaker movement, largely in and east of the Mississippi region, and Joan Baez Institute for the Study of Nonviolence on the West Coast, are the 2 principle groups conducting education in nonviolent revolution in this country.\\
The Peacemaker movement is secular. ``Catholic Worker" tells you what that is. But the Catholic worker is derived from \emph{Christianity}. It is a tribute to Catholic Workers that they cd derive an-archism from obedience training, a swamp of Christian sentimentality and a fog of theological bullshit. In fact the an-archism gets in as something that the pope allows to Catholics on a par with political doctrines that contradict Jesus. Jesus was not really wishy-washy, and he left no room for politics or popes, as where Mark reports:
\blockquote{$^{42}$ But Jesus called them \emph{to him}, and saith unto them, Ye know that they which are accounted to rule over the Gentiles exercise lordship over them and their great ones exercise authority upon them.\\
$^{43}$ But so shall it not be among you; (Mark 10:)}
That is as good as a definition of an-archism.\\
In terms of \emph{rule, lordship, authority}, Jesus accounted for the Government-thing, The State, and then he knocked it down flat. NO COPS $\;$ NO COURTS $\;$ NO JAILS $\;$ NO TAXES. FREEDOM NOW is the way Jesus said it again in Luke 17:20--21. And Jesus removed all obligation to pay tax in Matt. 17:24--26. And so on and so on...\\
The Catholic Worker movement is really based on the pope, not on Jesus. If the pope decided to put Catholic an-archism in the same bag with Bolshevism, it wd go out with the garbage tomorrow.\\
However, the Catholic Worker movement came in during the depression. it got started largely as a response to the hard times endemic to big cities in both good times and bad. They differentiate \emph{poverty} from \emph{destitution} and try to bring the worst deprivations up to their level, of voluntary poverty, with works of mercy, like soup kitchens, in their Houses of Hospitality. They --- some of them --- identify themselves as an-archists with an-archism, unlike the Peacemakers who have ambiguous relationships with the term. But as an-archists, Catholic Workers take much the same positions as do Peacemakers: draft card burning, tax-refusal, anti-racism, nonviolent revolution. It was the Cathlic Worker outfit, in NYC, which did much to defeat air raid drills and thereby raise consciousness in regard to nuclear war. However, Catholic Worker encouragement to draft- \& tax-law breaking is muted, as compared to Peacemakers. So it appears, at any rate, on the journalistic level. And journalistic write-ups of its an-archist position are infrequent and sometimes contradicted by reformist material, but they do produce books, among them the autobiografy of Ammon Hennacy (\emph{The Book of Ammon}) and \emph{The Long Loneliness} by Dorothy Day. They have Friday evening meetings for the clarification of thought in which I have participated in Chicago. They have an urban presence, as compared to Peacemakers who are diffused all over the place. Weak and uncertain as Catholic Worker propaganda for an-archism often is, sometimes it has been the only thing going for an-archism at all.\\
But there is on important difference between the Catholic Worker and the Peacemaker. Peacemakers support the position of nonviolent resistance to arrest which means \emph{don't walk when kidnappers or cops take and keep you prisoner}. This position extends into not standing for a judge, fasting in captivity and not working while in captivity. Some Catholic Workers, in their journalistic expressions, have been uncharitable to this ``going limp" as it is called. They are even uncharitable to the \emph{person} who went limp when arrested. ``Going limp" as a form of mutiny, seems to be a relatively new development in pacifist \& an-archist behavior. There seems to be no traditional heroes who have resisted Authority in this fashion, and since the Catholic Workers, as Catholics, follow heroes and leaders a bit more slavishly than they shd they do no know how to understand people like Corbet Bishop who went for well over 400 days without walking, eating, or taking a shit for the democratic taxpayers who made him a prisoner, but cd not make him a slave.\\
I have gone limp about a dozen times, almost hald the number of times I have been arrested which total 30 times. The last time I went limp under arrest was 16 Nov 74 when I was arrested by Patrolman Dennis Oppedisano \# 13503, because I was a black person on Lincoln Ave. He told me twice that he arrested me because I was a black on Lincoln Ave, and when he showed me his ITALIAN POWER button, inside his police jacket, he said that it meant that ``they" were going to kill all blacks in Illinois.

\section*{8}
I said that you can attack racism by attacking patriotism and I propose to do that with this poetry

\begin{Verbatim}[commandchars=+\[\]]
                    did St George slay the Indian?
          HOW THE BUY CENTENNIAL +underline[MUST] BE RACIST
          
          Shd we judge slaveowners by the standards they set?
It surely wd be to judge them by the standards of their time
                         rather
                    than by the subversions of Jesus
                    who belongs to no one's present...
                         the verse of Diderot
               or that Thomas who was a +underline[pain]
           wherever Jeffersons did variations
           on the themes    of Authority:
                               revolutions:
                that +underline[counter] Enlightenments
                     of Sade Roux Varlet...
                     et les Enrages: who were they?
                     
                          Judging History
                          By standards
                          Of the literate set
           decertifies    +underline[Her]   story
           and awards Respectability
      like the
           objectivity
                that takes little note
                of what was written on the wind
                by Africans working out
                in their own time
           except that a rhythm of liberty
                KILL THE WHITES
                KILL THE WHITES
      was suppressed by such pens
      as were to pen
      declarations of independence
           from FREEDOM
           
                +underline[NOW]
                   (and them...)
                   
 Was Jefferson greater than Garrison
Or did Jefferson rate because garrisons
 Condemn   Tolstoyans, Godwins and all god'schillun
 To that marginal textbook fate
   in whose context
     Centennial buyers
       do no read out
         catholic inquisitions
           into
             rela+underline[tiv]ity
               of standards
 
 But History be writ in blood:
      Workers in Haiti
           Adams' contemp'raries
      judging: judgers of monarchy
      by their own Rule
           found them wanting
           and found necessary
           to slaughter them
                so t+underline[HA]t +underline[I]sland's space migh+underline[T] approx+underline[I]mate
                dead +underline[masters']
                ideals
                ...which
                     ...strange to say
                     ...it does to this day...
                the gui llo tine will voodoo you
                the gui llo tine will voodoo you
                
           But come now
                let's capture the spirit of Concord
                in this ameriKKKan obscene
                obscen   ario
                     for
                     July 4
                197 sick sick six:
                     
           +underline[black woman]
           pregnant kneeling shackled
           is reading   declaration of no dependence
           
                +underline[patriot]
                     standing
                          lashes her black
                          in time to +underline[Stars & Stripes Forever],
                          Chiming Liberty Bell on upswing
                     
           When it gets down to / pledging
                Fortunes, Lives and sacred Honor
                     He rapes her.
                          He
                     Climaxes
                   on
                     Nukey Nukey
                     Vietnam
                     Nixon and the atomBOMB
                Kingdom Come is Einstein-way...
                To Holocaust the world away.  But
                
           this poem is not compleat
                until you
                     match the flag
                     with an-archist
                          +underline[fire].
\end{Verbatim}

\chapter{Why I Am Not an Anarchist}
\chapterauthor{Gregory Hill}

About five years ago I considered myself an anarchist (anarchopacifist, in particular), because I believe that the highest authority available to any individual is one's own honest experience and that any other authority provides only vicarious information at best.\\
I've not changed my opinion about this, but I have ceased referring to myself as an anarchist. The reason is basic and simple: TOO DAMN MANY RULES.\\
OK, it's a joke. But it's a TRUE joke. The incompatibility is not between my position and some anarchist theories, but between my position and the position of most of those who use the label ``anarchist."\\
It seems that Rule Number One of anarchy, as understood by authoritarians and by most who call themselves anarchists, is that a government is an enemy. Rule Number Two is that to gain freedom the individual is politically or morally or somehow obligated to fight this enemy.\\
In my opinion, these rules represent a position which would be better referred to as anti-archy. The prefix ``a" means ``without" and it need not imply ``against". There is an exact parallel with the word atheist --- it is usually used and understood, by those for it and against it, as though the word was anti-theist.\\
I can respect the anti-archist position, but I don't share it. The government is not my enemy because there is no government. OK, another joke, but still a TRUE joke. I know good and well that there are people with guns who restrict my free decisions, and I know about groups of people collecting taxes from me, and all the rest of this government business perceive it in the same manner that I perceive (for example) a big rock in my path which necessitates stepping around and compromising myself. Frankly, I don't believe in rocks either --- I just step around and compromise (which is actually easier than is believing in them). I think that there is a big difference in gree between (a) existentially responding to a phenomenon and (b) conceptualizing it as an ``enemy." If everything in the universe that has ever thwarted my purpose is my enemy, the only nothing can be my friend --- and that excludes even myself. But, still, I respect the anti-archist position. After all, if one does perceive a phenomenon to be an enemy then one would be a damn fool to do other than defend oneself.\\
Much of this essay is futzing around with labels. Still, I feel free to futz, and in any case what I'm trying to do is to avoid the assumption by others that I am at war with certain people just because those people think that they are a government and go out of their way to forcibly impose their notions on me.\\
I'm not at war with them or with rocks either. And insofar as anyone thinks that an anarchist is one who is supposed to do something or another, the there are too damn many rules for me ant to hell with the whole business.

\chapter{The Ether Vibrates}
\vspace{-1cm}
\emph{Letters from Readers}\\

\noindent \textsc{Wantagh, Long Island, New York:}\\

It reassured me, after I read Robert Shea's article in \emph{No Governor} \# 1, that the effective anarchist group doesn't consist of those proximate to you in space or convenience of access, but those proximate in purpose, energy and initiative. For the rest, ``\emph{all} contributions gladly accepted," especially including constructive criticism. But know thyself --- don't promise more than you \emph{know} you will deliver. Modern communications certainly ought to be able to help the active ones coordinate.
\par\begin{flushright} --- Robert Braine\\ \end{flushright}

\noindent Ed Note: We received a very long letter from Joffre Stewart which we're saving for the next issue. We'd like to receive and print more letters from readers.




\chapter{The Frying Pan}
\vspace{-1cm}
\emph{Reviews of Libertarian Publications Recently Received}\\

\noindent\textsc{Against the Wall} A magazine of self-liberation and voluntary alternatives. Volume 4, Number 1 includes a short story that takes place during the Bolshevik siege if free Kronstadt and an article by Robert Anton Wilson on the philosophy of Timothy Leary. A sample issue is 75 cents postpaid. Against the Wall, P.O. Box 444, Westfield, New Jersey 07091.\\

\noindent\textsc{Black Star} An Anarchist Review. A publication of the Social Revolutionary Anarchist Federation. Volume 1, Number 1 includes ``Some Thoughts on Anarchist Culture" by Glenn Meredith; an article on how people's efforts prevented a university from taking over and closing a public park, by Louise Crowley; ``Rape, a Crime?" by Jim Bumpas; and much more, including poems, drawings and cartoons. 5 cents a copy, 6 issues for \$3 (the higher subscription price is intended to help defray production costs). The second issue should be available any time now. Black Star, Box 92246, Milwaukee, Wisconsin 53202.\\

\noindent\textsc{Black Flag of Anarchism} A one-page flier of news, quotations and graphics. Printed occasionally by Solidarity Collective, RWC 3 south rm 346, Bristol, Rhode Island 02809. Donations requested.\\

\noindent\textsc{Libero International} An anarchist magazine from Japan written in English for an international audience. Published by a federation of autonomous libertarian groups called CIRA-Nippon, edited by the federation's Section for International Correspondence. The second issue, dated May 1975 has articles on anarchist organizations in Japan, hospital conditions, the Korean anarchist movement and the history of anarchism in Asia. Subscriptions \$3 a year (six issues), single copies 50 cents. Send money orders or cash, not personal checks. Libero International, CPO Box 1065, Kobe, Japan 650--91.\\

\noindent\textsc{Solidarity Newsletter} National and international news of interest to anarchists. 15 cents a copy, ten issues for \$1.50. Philadelphia Solidarity, GPO Box 13011, Philadelphia, Pennsylvania 19101.\\

\noindent\textsc{Sweet Gherkins} from the Dill Pickle Press. October 24 issue includes excerpts from the writing of Franklin P. Adams, Julia Peterkin, Wilhelm Reich, J. David Stern, Joseph Joubert, Vincent Sheean and Marcus Aurelius Antoninus. There are also mini-articles on behaviorism and fascism by the editor, Edward R. Johnson. P.O. Box 822, Muscatine, Iowa 52761.\\

\noindent Just before we went to press we got these publications:\\

\noindent\textsc{Against the Wall} Volume 4 Number 4. This ``End of the World Issue" could really be the end of this publication. We suspect a lot of readers will want to put the publishers up against the wall. 50 cents. P.O. Box 444, Westfield, New Jersey 07091.\\

\noindent\textsc{Other Scenes} The International Newspaper. We were at the annual conference of the National Organization for the Reform of Marijuana Laws last December and there John Wilcock, who handed us a copy of his October 1975 issue. He has been producing the paper in different forms in different countries since 1966. Subscriptions \$10 a year. BCM-OSCENES, London WC1V, England.\\

\noindent\textsc{Black Star} An Anarchist Review. Volume 1, Number 2 is an enormous improvement, graphically and every other way, over the first issue of this general propaganda organ of the Social Revolutionary Anarchist Federation. Includes articles on rape, intellectuals, the 1975 Hunter College conference of anarchists and libertarians, workers' control and the prostitutes' revolt in France. The new cover price is 25 cents ``but take it if you're broke." P.O. Box 92246, Milwaukee, Wisconsin 53202.\\

\noindent\textsc{Sweet Gherkins} from the Bill Pickle Press. November 24 issue has a beautiful haiku by Basho, excerpts from James Gibbons Huneker, Julia Ward Howe, Ilya Ehrenburg and Herman Melville, among others. Articles by ERJ on fascism and behaviorism. Subscription 10 issues for \$1, single copy 10 cents.



\end{document}


















