\chapter{The Simple Life}
\chapterauthor{Joffre Stewart}

\section*{Prefatory note}
Most (practically all) of the material below was delivered in a class on the Simple Life, taught by one Bill Ray at Francis W. Parker School (Chicago), on TUE 20 MAY 75. Out of 12 students, 2 helped to complete poem at end by putting match to wee flag. The student publication aired some debate regarding the burning of the Constitution, etc.

\section*{1}
Presenting an-archism from the angle of the Simple Life, as I do here, will have 2 limitations. It will not develop the an-archist critique of Marxism. Marxists-in-power have been the most dangerous enemies of anti-statism, of an-archism, since 1917. The 2nd limitation is that little will be said of the world-wife an-archist movement as it exists today.

\section*{2}
Some an-archists, called \emph{primitive an-archists}, have turned to anthropology and turned their attention toward certain nonliterate societies as the \emph{natural} societies that best take human nature into account. They want to pattern their an-archy on the best of such societies.\\
Partly because of this, the Simple Life may be seen as an an-archistic criticism of civilization. Or, a criticism of civilization, leveled at it by the an-archist spirit. It does not matter what kind of civilization, all it has to be is a civilization. This is because too many things that a wise person in the simple life can do without, are things collected by foul means, by car, exploitation, slavery, coercion. Therefore it happens, quite logically, that the simple life appears as an expression of pacifist an-archism. An' sho 'nuff: the first an-archists who left a record were pacifist an-archists: Lao-Tzu, the Greek Cynics \& Stoics, and Jesus of Nazareth. So the first an-archist, anarcho-pacifist, was a Chinaperson, Lao-Tzu, who, after expressing his way of life in the poetry called the Tao Te Ching, is said to have quit his job as a government burocrat and went off to the mountains to live the simplest life.\\
The Cynics, and the Stoic, Zeno of Citium who learned from the Cynics, tended to come from the lowest class in Greek society that had enuf liberty \& leisure to filosofize. That is, they were a cut above the slaves, barely out of slavery. The point is that their filosofy was a rational expression of their class interest because they were against slavery. They understood that to be against slavery you had to be against was because slaves are prisoners of war. \emph{Peace \& Freedom} are one and the same thing. The person who tries to sell you one without the other is a politician and the politician is the greatest danger to both.\\
The Cynics are renowned for their simplicity in a civilization that was pretty uncomplicated as it was compared to the Sears Towers, Apollo space ships, TV, computer spying, long forms, legal jargon, traffic jams, wiretaps, ice cream cones and radiation diseases of the present civilization. But on of these Cynics lived in a tub. When one of them, perhaps the same one, saw a boy drink water with his hands, he threw away his cup. When Alexander the Great asked the wise man what cd he do for him, the wise man asked the Great Man to step aside so he cd get some sun.\\
Jesus was nonviolent, and because he understood his position, he was not a liberal, he was an-archist, he was a pacifist revolutionary. The Son of Man, as Jesus called himself --- Jesus practically never called himself `Christ' and there is an an-archist reason for this --- lived such a simple life that he had nowhere to lay his head. He told the rich people to give it all away and follow him. He said that it wd be easier for a camel to go thru the eye of a needle than for a rich man to enter the kingdom of heaven. By ``Kingdom of Heaven" Jesus meant an-archy, paradise regained. Tho Jesus worked as a carpenter, he contradicts the Protestant ethic where he asks you to dig on the lilies of the field: they neither spin nor sew, and yet, blooming naked, they got it all: Solomon is all his glory cd not compete. Tho we know less about the sex life of Jesus that we know about god's --- or rather because of that --- it is not necessary to interpret the Nazarene version of the simple life the way the priesthood has. Jesus indicated that he came to encourage abundant living and abundant living can mean all of sexual fulfillment.

\section*{3}
The facts about these people in the classical tradition are usually taught, if at all, in a superficial way that ignores and covers-up the fact that they were against politics, Governments, wars, and bosses. But a lot of schoolteaching time lay be spent on slave owners like Aristotle or totalitarians like Plato. Jesus belongs to classical times but Jesus is not taught because he is controversial, but however, controversial for the wrong reasons, the wrong reasons being like those unimportant reasons that separate Protestant from Catholic. Jesus is no controversial for the right reasons: the an-archist teachings which caused the Protestants \& Catholics to gang up on the people who \emph{did} correctly understand Jesus and destroy them in 1535. Those people were called Anabaptists. And to this day, the survivors of those 16th Century Anabaptists, called Amish, Brethren, Mennonites, Quakers, may prefer the horse to farm machinery, will not address persons by titles, will not send children to secondary school and in other ways show a preference for the simple life, whether you agree with the details or not.\\
Christianity, its divinity schools \& complicated theologies exist in order to suppress the important truths about Jesus; and they exist because Jesus' anti-Statism is suppressed. To know and live the simple truths of Jesus is to destroy Christian civilization.

\section*{4}
The an-archists of modern times who are attracted to simplification as part of the solution also bring nonviolence with them as part of the solution.\\
A graduate of Harvard named Henry David Thoreau, who wrote an \emph{Essay on Resistance to Civil Government}, does not fit exactly into either the an-archist or pacifist category. But he was an anti-slavery man and \emph{Freedom is what an-archism is all about.} Freedom is what an-archism is all about, so as a LIBERTARIAN ANTI-SLAVERY INDIVIDUALIST, Thoreau was very close to an-archism, and by practicing this \emph{near}-anarchism, he has taught the modern world as much about behaving non-violently as has Jesus. And, as with the Cynics, Stoics, and other ancients, this convergence of nonviolence and anti-Statism leads into the simple life.\\
For on thing, the simple life was necessary for Thoreau because he did not want to pay taxes. TAXATION IS ROBBERY. The person who taxpays regularly is in slavery. In order to not pay taxes, Thoreau lived on the margins of the economy in such ways that the tax-collector cd not collect.\\
And I live where Thoreau lived: on the margins of the economy. I have not made a payroll since August 1943, therefore, I have not rendered payroll tax, \emph{income tax}, since 1943. I did not pay for the Korean war, the Viet Nam war, and I ma not paying to keep those people in jail who resisted the Viet Nam war. I did not pay to kill the Rosenbergs. I did not pay the Bay of Pigs invasion, and I did not pay the CIA to kill Malcolm X, JFK, J. Edgar Hoover, and 1,000,000 Indonesians in 1965. I am not paying Ford-Kissinger to send Marines to commit violations in Cambodia. I did not buy the 2 atomic BOMBS that Nixon wanted to give the French to dump on the Vietnamese. I did not pay Nixon's salary. The taxpayer may be the biggest criminal you can find, and you shd consider it an honor to be persecuted for o p e n l y refusing to pay taxes if ever it gets around to that.\\
The Russian Count, Leo Tolstoy, reckoned that you don't fight against war by going to war, and he recognized The State as a (very necessary) cause of war and so became an an-archist altho he did not wear the label. And like an-archist pacifists since Lao-Tzu 2500 years ago, his beliefs took him into the practice of the simple life. He renounced all the fashionable ideas of aristocratic, ruling class life, and began to live the simple life of the peasants on his estates, and began to do constructive things that helped out this simple life.\\
Tolstoy recognized that he did not belong to the power arrangements that arrange for masses of people to kill each other in war, and because he was not part of those power arrangements, he was therefore not a citizen of Russia of of any other country. He was a STATELESS PERSON OF THE WORLD, much life the Stoic, Zeno of Citium, thousands of years before him. It might also be said that Thoreau was not a citizen of the United States or of any other country, but Thoreau had a more sly way of saying it which was self-contradictory as well.\\
I renounced citizenship around Nagasaki Day 1950, declaring myself A STATELESS PERSON OF THE WORLD. The point is that the pacifist an-archists, inclined toward the simple life, tend to have a better idea of who they are in regard to the claims of The State on their identity. The an-archists who are coming from and who are closer to a labor and socialist background are less clear about who they are and this can be a weakness when it comes to refusing to be drafted. The proletarian class concept tends to deal with a part of a person's humanity rather than all of it. This concept has seen the grossest failure in 2 world wars. Even Kropotkin forgot that he was anti-capitalist and took sides in WWI. It is easier for anyone who is not a pacifist to slide into any kind of war that for someone who is pacifist. In this day and age, the an-archism which is not uncompromisingly pacifist is especially weak and irrelevant.\\
In this regard, it is impossible to overemfasize the desirability of attacking and destroying patriotism. One of the biggest failures o an-archism today is that it is not shooting down patriotism. Patriotism is more dangerous than racism, altho the racism that drove so many Jews into Zionism may destroy the world. On the other hand, one way to attack racism is by attacking patriotism, and I will try to show what I mean later...

\section*{5}
Right now I will deal with the Equal Rights Amendment. The Equal Rights Amendment will have women thrown in jail for not carrying draft cards and more than double the amount of cannon fodder available to the Pentagon. I don't believe in jails, I don't believe in war, and an-archism is against all the Law that organizes jail \& war. So what does that mean for the Constitution?
\begin{center}
(\emph{burned copy of Constitution})
\end{center}
By the light of this flaming Constitution, you see plainly why an-archists can never think of going for the Equal Rights Amendment. If there is no Constitution, there is no place to put an amendment. Liberating society from the State liberates women from a male dominated institution that inherently puts women at a disadvantage. An-archism is feminism's best bet. And without the State, there is nothing to be patriotic about. It is as simple as that.

\section*{6}
This brings us to present time. The 2 an-archists best known to Americans are Joan Baez and John Lennon. This is an ambiguous distinction because they are better known as singers than as an-archists. And being so well known interferes with their an-archism because they are rich and this is too fancy for some an-archists from the labor, libertarian socialist background: they tend to discriminate against John \& Joan on account of their ``class." I think class discrimination is unfortunate no matter who does it to whom. Rather than ignore John \& Joan as an-archists, as some IWW types do, I think it is better to acknowledge them as an-archists, and relate to them by criticizing their shortcomings. That way everybody benefits.\\
The people criticize Joan Baez for her money, yet have a lot to learn from her: she pulls down the flag and cuts it to pieces with her tongue. She gives patriotism the trouncing it deserves. She obstructed an induction center with her body. And her money helped reprint some an-archist classics.\\
But she is insincere and incomplete when it comes to tax-refusal: she makes a gesture of refusal but leaves money in the bank where the government can seize it. This is what she shd be criticized for, not merely for having the money. If the pacifist Joan Baez were for real as a pacifist tax-refuser, she would find it very difficult to get or keep money and her life style might be more simple than I think it is because she might not have a choice. My understanding is that her husband left her because her life style reflected her high earnings...\\
I don't know the calculation on John Lennon's personal appointments: the image I have is that they are modest even tho he's rich. What is worse, however, is that he is not getting in trouble for conscientiously refusing taxes.\\
But he and the other Beatles, for an-archist reasons, did reject the Queen's honors. This means a lot more than when Viet Nam vets throw their medals back at the Government. He is not Sir John Lennon and he cd be. The Beatles also tried to destroy fashion, but they were trapped in their leadership and only succeeded in setting new fashions. There is a difference between being a fan/a follower and being an an-archist.\\
But John Lennon at least acknowledges in more ways than one that the simple life is the direction in which he ought to go. Not only does he tell us this in his song writing, but he associates the simple life with out-and-out atheist an-archism. The title of they lyrics is \emph{Imagine}: 
\begin{center}
\begin{tabular}{l l l l}
Imagine there's no heaven & & &  imagine no possessions\\
it's easy if you try & & & I wonder if you can\\
no hell below us & & &  no need for greed of hunger\\
above us only sky & & &  $\;\;$ a brotherhood of man\\
imagine all the people & & &  imagine all the people\\
living for today... & & & sharing all the world...\\
\\
imagine there's no countries & & & you may say I'm a dreamer\\
it isn't hard to do & & & but I'm not the only one\\
nothing to kill or die for & & & I hope someday you'll join us\\
and no religion too & & & and the world will be as one\\
imagine all the people\\
living life in peace...
\end{tabular}
\end{center}
John Lennon is up for deportation because he is such a good poet. AmeriKKKa does not want the best and tries to get rid of it in any way convenient. But the an-archist movement is weirdly lacking in solidarity whenever it cannot rally behind such a case as his.

\section*{7}
The Peacemaker movement has developed the notion of simple living into one of its active principles. The Catholic Worker movement lays even more stress on what it calls \emph{voluntary poverty}. Both movements are an-archist, both are pacifist, and relations between them are cooperative and mutually reinforcing. In terms of the circulation of their periodicals (\emph{The Peachemaker}; \emph{The Catholic Worker}) these two account form by far the bulk of the an-archist movement in ameriKKKa (but this way of reckoning can be tricky...) In terms of lasting power and continuity, these 2 \emph{are} the an-archist movement in ameriKKKa. But in terms of theoretical discussion, their activity is minimal, essentially unhistorical, and unsatisfying for people who have wide-ranging needs to know. What they do do is the propaganda of the deed: the nonviolent ACT.\\
The Peacemaker movement began around 1948. It came into existence to break draft laws, especially to encourage \emph{you} not to register for the draft. Nor do its members do war work. By 1952, the nonpayment of taxes was as firmly a part of Peacemakers as refusal to register for the draft. From the beginning, Peacemakers adopted the notion of inner personality transformation and of nonviolent revolution but the idea of nonviolent revolution did not clearly take on an-archist characteristics until sometime between 1962 and 1972. THE IDEA OF SIMPLE LIVING MUST HAVE BEEN THERE ALL THE TIME, but the recognition of it as one of their characteristic properties and practices came slowly and built up gradually in time. A column on Simple Living often appears in the Peacemaker publication.\\
The Peacemaker movement, largely in and east of the Mississippi region, and Joan Baez Institute for the Study of Nonviolence on the West Coast, are the 2 principle groups conducting education in nonviolent revolution in this country.\\
The Peacemaker movement is secular. ``Catholic Worker" tells you what that is. But the Catholic worker is derived from \emph{Christianity}. It is a tribute to Catholic Workers that they cd derive an-archism from obedience training, a swamp of Christian sentimentality and a fog of theological bullshit. In fact the an-archism gets in as something that the pope allows to Catholics on a par with political doctrines that contradict Jesus. Jesus was not really wishy-washy, and he left no room for politics or popes, as where Mark reports:
\blockquote{$^{42}$ But Jesus called them \emph{to him}, and saith unto them, Ye know that they which are accounted to rule over the Gentiles exercise lordship over them and their great ones exercise authority upon them.\\
$^{43}$ But so shall it not be among you; (Mark 10:)}
That is as good as a definition of an-archism.\\
In terms of \emph{rule, lordship, authority}, Jesus accounted for the Government-thing, The State, and then he knocked it down flat. NO COPS $\;$ NO COURTS $\;$ NO JAILS $\;$ NO TAXES. FREEDOM NOW is the way Jesus said it again in Luke 17:20--21. And Jesus removed all obligation to pay tax in Matt. 17:24--26. And so on and so on...\\
The Catholic Worker movement is really based on the pope, not on Jesus. If the pope decided to put Catholic an-archism in the same bag with Bolshevism, it wd go out with the garbage tomorrow.\\
However, the Catholic Worker movement came in during the depression. it got started largely as a response to the hard times endemic to big cities in both good times and bad. They differentiate \emph{poverty} from \emph{destitution} and try to bring the worst deprivations up to their level, of voluntary poverty, with works of mercy, like soup kitchens, in their Houses of Hospitality. They --- some of them --- identify themselves as an-archists with an-archism, unlike the Peacemakers who have ambiguous relationships with the term. But as an-archists, Catholic Workers take much the same positions as do Peacemakers: draft card burning, tax-refusal, anti-racism, nonviolent revolution. It was the Cathlic Worker outfit, in NYC, which did much to defeat air raid drills and thereby raise consciousness in regard to nuclear war. However, Catholic Worker encouragement to draft- \& tax-law breaking is muted, as compared to Peacemakers. So it appears, at any rate, on the journalistic level. And journalistic write-ups of its an-archist position are infrequent and sometimes contradicted by reformist material, but they do produce books, among them the autobiografy of Ammon Hennacy (\emph{The Book of Ammon}) and \emph{The Long Loneliness} by Dorothy Day. They have Friday evening meetings for the clarification of thought in which I have participated in Chicago. They have an urban presence, as compared to Peacemakers who are diffused all over the place. Weak and uncertain as Catholic Worker propaganda for an-archism often is, sometimes it has been the only thing going for an-archism at all.\\
But there is on important difference between the Catholic Worker and the Peacemaker. Peacemakers support the position of nonviolent resistance to arrest which means \emph{don't walk when kidnappers or cops take and keep you prisoner}. This position extends into not standing for a judge, fasting in captivity and not working while in captivity. Some Catholic Workers, in their journalistic expressions, have been uncharitable to this ``going limp" as it is called. They are even uncharitable to the \emph{person} who went limp when arrested. ``Going limp" as a form of mutiny, seems to be a relatively new development in pacifist \& an-archist behavior. There seems to be no traditional heroes who have resisted Authority in this fashion, and since the Catholic Workers, as Catholics, follow heroes and leaders a bit more slavishly than they shd they do no know how to understand people like Corbet Bishop who went for well over 400 days without walking, eating, or taking a shit for the democratic taxpayers who made him a prisoner, but cd not make him a slave.\\
I have gone limp about a dozen times, almost hald the number of times I have been arrested which total 30 times. The last time I went limp under arrest was 16 Nov 74 when I was arrested by Patrolman Dennis Oppedisano \# 13503, because I was a black person on Lincoln Ave. He told me twice that he arrested me because I was a black on Lincoln Ave, and when he showed me his ITALIAN POWER button, inside his police jacket, he said that it meant that ``they" were going to kill all blacks in Illinois.

\section*{8}
I said that you can attack racism by attacking patriotism and I propose to do that with this poetry

\begin{Verbatim}[commandchars=+\[\]]
                    did St George slay the Indian?
          HOW THE BUY CENTENNIAL +underline[MUST] BE RACIST
          
          Shd we judge slaveowners by the standards they set?
It surely wd be to judge them by the standards of their time
                         rather
                    than by the subversions of Jesus
                    who belongs to no one's present...
                         the verse of Diderot
               or that Thomas who was a +underline[pain]
           wherever Jeffersons did variations
           on the themes    of Authority:
                               revolutions:
                that +underline[counter] Enlightenments
                     of Sade Roux Varlet...
                     et les Enrages: who were they?
                     
                          Judging History
                          By standards
                          Of the literate set
           decertifies    +underline[Her]   story
           and awards Respectability
      like the
           objectivity
                that takes little note
                of what was written on the wind
                by Africans working out
                in their own time
           except that a rhythm of liberty
                KILL THE WHITES
                KILL THE WHITES
      was suppressed by such pens
      as were to pen
      declarations of independence
           from FREEDOM
           
                +underline[NOW]
                   (and them...)
                   
 Was Jefferson greater than Garrison
Or did Jefferson rate because garrisons
 Condemn   Tolstoyans, Godwins and all god'schillun
 To that marginal textbook fate
   in whose context
     Centennial buyers
       do no read out
         catholic inquisitions
           into
             rela+underline[tiv]ity
               of standards
 
 But History be writ in blood:
      Workers in Haiti
           Adams' contemp'raries
      judging: judgers of monarchy
      by their own Rule
           found them wanting
           and found necessary
           to slaughter them
                so t+underline[HA]t +underline[I]sland's space migh+underline[T] approx+underline[I]mate
                dead +underline[masters']
                ideals
                ...which
                     ...strange to say
                     ...it does to this day...
                the gui llo tine will voodoo you
                the gui llo tine will voodoo you
                
           But come now
                let's capture the spirit of Concord
                in this ameriKKKan obscene
                obscen   ario
                     for
                     July 4
                197 sick sick six:
                     
           +underline[black woman]
           pregnant kneeling shackled
           is reading   declaration of no dependence
           
                +underline[patriot]
                     standing
                          lashes her black
                          in time to +underline[Stars & Stripes Forever],
                          Chiming Liberty Bell on upswing
                     
           When it gets down to / pledging
                Fortunes, Lives and sacred Honor
                     He rapes her.
                          He
                     Climaxes
                   on
                     Nukey Nukey
                     Vietnam
                     Nixon and the atomBOMB
                Kingdom Come is Einstein-way...
                To Holocaust the world away.  But
                
           this poem is not compleat
                until you
                     match the flag
                     with an-archist
                          +underline[fire].
\end{Verbatim}