\chapter{The Frying Pan}
\vspace{-1cm}
\emph{Reviews of Libertarian Publications Recently Received}\\

\noindent\textsc{Against the Wall} A magazine of self-liberation and voluntary alternatives. Volume 4, Number 1 includes a short story that takes place during the Bolshevik siege if free Kronstadt and an article by Robert Anton Wilson on the philosophy of Timothy Leary. A sample issue is 75 cents postpaid. Against the Wall, P.O. Box 444, Westfield, New Jersey 07091.\\

\noindent\textsc{Black Star} An Anarchist Review. A publication of the Social Revolutionary Anarchist Federation. Volume 1, Number 1 includes ``Some Thoughts on Anarchist Culture" by Glenn Meredith; an article on how people's efforts prevented a university from taking over and closing a public park, by Louise Crowley; ``Rape, a Crime?" by Jim Bumpas; and much more, including poems, drawings and cartoons. 5 cents a copy, 6 issues for \$3 (the higher subscription price is intended to help defray production costs). The second issue should be available any time now. Black Star, Box 92246, Milwaukee, Wisconsin 53202.\\

\noindent\textsc{Black Flag of Anarchism} A one-page flier of news, quotations and graphics. Printed occasionally by Solidarity Collective, RWC 3 south rm 346, Bristol, Rhode Island 02809. Donations requested.\\

\noindent\textsc{Libero International} An anarchist magazine from Japan written in English for an international audience. Published by a federation of autonomous libertarian groups called CIRA-Nippon, edited by the federation's Section for International Correspondence. The second issue, dated May 1975 has articles on anarchist organizations in Japan, hospital conditions, the Korean anarchist movement and the history of anarchism in Asia. Subscriptions \$3 a year (six issues), single copies 50 cents. Send money orders or cash, not personal checks. Libero International, CPO Box 1065, Kobe, Japan 650--91.\\

\noindent\textsc{Solidarity Newsletter} National and international news of interest to anarchists. 15 cents a copy, ten issues for \$1.50. Philadelphia Solidarity, GPO Box 13011, Philadelphia, Pennsylvania 19101.\\

\noindent\textsc{Sweet Gherkins} from the Dill Pickle Press. October 24 issue includes excerpts from the writing of Franklin P. Adams, Julia Peterkin, Wilhelm Reich, J. David Stern, Joseph Joubert, Vincent Sheean and Marcus Aurelius Antoninus. There are also mini-articles on behaviorism and fascism by the editor, Edward R. Johnson. P.O. Box 822, Muscatine, Iowa 52761.\\

\noindent Just before we went to press we got these publications:\\

\noindent\textsc{Against the Wall} Volume 4 Number 4. This ``End of the World Issue" could really be the end of this publication. We suspect a lot of readers will want to put the publishers up against the wall. 50 cents. P.O. Box 444, Westfield, New Jersey 07091.\\

\noindent\textsc{Other Scenes} The International Newspaper. We were at the annual conference of the National Organization for the Reform of Marijuana Laws last December and there John Wilcock, who handed us a copy of his October 1975 issue. He has been producing the paper in different forms in different countries since 1966. Subscriptions \$10 a year. BCM-OSCENES, London WC1V, England.\\

\noindent\textsc{Black Star} An Anarchist Review. Volume 1, Number 2 is an enormous improvement, graphically and every other way, over the first issue of this general propaganda organ of the Social Revolutionary Anarchist Federation. Includes articles on rape, intellectuals, the 1975 Hunter College conference of anarchists and libertarians, workers' control and the prostitutes' revolt in France. The new cover price is 25 cents ``but take it if you're broke." P.O. Box 92246, Milwaukee, Wisconsin 53202.\\

\noindent\textsc{Sweet Gherkins} from the Bill Pickle Press. November 24 issue has a beautiful haiku by Basho, excerpts from James Gibbons Huneker, Julia Ward Howe, Ilya Ehrenburg and Herman Melville, among others. Articles by ERJ on fascism and behaviorism. Subscription 10 issues for \$1, single copy 10 cents.
